\documentclass{book}


% Set paper and layout
%
\usepackage[landscape, top=1cm, left=1cm, right=1cm, bottom=1cm]{geometry}
\setlength\parskip{-0.05cm} % no extra vertical space between paragraphs
\setlength\parindent{0cm} % no indent on first lines of paragraphs
\linespread{1.75} % vertical line-spacing and vertical paragraph spacing


% Use Paracol package to do two-column format
%
\usepackage[english]{babel} % English as "system" language, paracol demands a specification
\usepackage{paracol} % \begin{paracol}{...} ... \switchcolumn[...] ... \end{paracol}
\setlength\columnsep{1.5cm} % seperation between columns
\newcommand{\s}{\switchcolumn} % a shorter command to replace \switchcolumn[...] from the paracol-package


% Use Endnotes packages, so ``footnotes'' go at end of document rather than the foot of each page
%
\usepackage{endnotes} % \endnote{...}
\addtoendnotes{\vspace*{-1em}} % no extra vertical space between the endnotes-header and the endnotes
\input{hyperendnotes.sty} % LOADS hyperendnotes.sty
\def\enoteheading{} % to have endnotes without the usual heading


% Calc package is necessary.  Does some math on lengths?  
%
\usepackage{calc}


% Use Graphicx package to rotate or flip text
% (eg. \reflectbox{C} --- \rotatebox[origin=c]{180}{V} --- \reflectbox{D})
%
\usepackage{graphicx} 


% Use math packages for fonts, symbols, etc.
%
\usepackage{fouriernc} % for $\otherpi$
\usepackage{upgreek} % certain upright greek characters
\usepackage{amsmath} % for $\text{cl.}^{\text{mus}}$
\usepackage{amssymb} % to have standard symbols for special sets --- to have $\nless$

\newcommand{\dittoMarkLatin}{\guillemotright}
\newcommand{\dittoMarkEnglish}{''}

% Peano's reverse C, that is used for implication.
\newcommand{\C}{\mathop{\sbox0{$\displaystyle$$\reflectbox{C}$} % reflect box mirrors upside-down and left-to-right
\raisebox{-\height+\ht0-0.05cm}[\ht0][\dp0]{\scalebox{0.8}[0.8]{\copy0}}}\displaylimits} %
% Peano's one dot, used to indicate precidence in equations
\newcommand{\p}{\mathop{\sbox0{$\displaystyle$$.$}
\raisebox{-\height+\ht0}[\ht0][\dp0]{\scalebox{1}[1]{\copy0}}}\displaylimits} % creates a "math"-point to have equal spacing
% Peano's 2 dots is just a colon :, and is used to indicate precidence in equations
% Peano's 3 dots, used to indicate precidence in equations
\newcommand{\pppNoSpace}{\leavevmode\lower0ex\hbox{.}\kern-0.1em\raise0.7ex\hbox{.}\kern-0.1em\lower0ex\hbox{.}} % arranging three points in a triangle
\newcommand{\ppp}{\thinspace \pppNoSpace\thinspace} % arranging three points in a triangle

\newcommand{\no}{\mathop{\thinspace \scalebox{1.5}{-}}}
\newcommand{\abs}{\mathop{\rotatebox[origin=c]{180}{V}}}
\newcommand{\mini}{\mathop{\rotatebox[origin=c]{180}{M}}}
\newcommand{\larger}{\mathop{\thinspace > \thinspace}}
\newcommand{\smaller}{\mathop{\thinspace < \thinspace}}
\newcommand{\such}{\rotatebox[origin=c]{180}{$\epsilon \thinspace$}}
\newcommand{\mult}{\mathop{\reflectbox{D}}}

\newcommand{\I}{\mathop{\thinspace \textbf{I} \thinspace}}
\newcommand{\E}{\mathop{\thinspace \textbf{E} \thinspace}}
\newcommand{\Lfat}{\mathop{\thinspace \textbf{L} \thinspace}}
\newcommand{\D}{\mathop{\thinspace \text{D} \thinspace}}
\newcommand{\K}{\mathop{\thinspace \text{K} \thinspace}}
\newcommand{\R}{\mathop{\thinspace \text{R} \thinspace}}
\newcommand{\M}{\mathop{\thinspace \text{M} \thinspace}}
\newcommand{\N}{\mathop{\thinspace \text{N} \thinspace}}
\newcommand{\Np}{\mathop{\thinspace \text{Np} \thinspace}}
\newcommand{\Q}{\mathop{\thinspace \text{Q} \thinspace}}
\newcommand{\T}{\mathop{\thinspace \text{T} \thinspace}}

\newcommand{\II}{\mathop{\thinspace \textbf{I\hspace{0.02cm}I} \thinspace}}
\newcommand{\EE}{\mathop{\thinspace \textbf{E\hspace{0.02cm}E} \thinspace}}
\newcommand{\LfatLfat}{\mathop{\thinspace \textbf{L\hspace{0.02cm}L} \thinspace}}



% Use Dashrule package for dashed lines (using \hdashrule)
%
\usepackage{dashrule} 


% Use Xcolor package to grey out unimportant text, etc.
%
\usepackage{xcolor}

% na for not applicable
% (these pieces of text are not important for mathematical
%  understand when reading the english text)
\newcommand\na[1]{\textcolor{gray}{#1}} 
% fn for footnote (notations / text which ask for footnote-like attention)
\newcommand\fn[1]{\textcolor{red}{#1}} 

\definecolor{blue}{RGB}{89,141,201}
\definecolor{blue_edge}{RGB}{53,126,189}

\definecolor{grey}{RGB}{235,235,235}
\definecolor{grey_edge}{RGB}{173,173,173}


% Use Hyperref package for URLs
%   ShareLaTeX.com said ``Usually [hyperref] has to be the last package to be imported''
%
\usepackage[colorlinks=true,linkcolor=red, urlcolor=red]{hyperref} % color links and url's --- hyperref for hyperlinks in endnotes
\urlstyle{same} % standard font for url's


% Heading commands
%    - assumes it is already in a centered format.
%
\newcommand\peanoHeadingSmall[1]{ \vspace{0.75cm} \textit{#1} \nopagebreak[4]

\vspace{0.25cm} \nopagebreak[1]}
\newcommand\peanoHeadingMedium[1]{ \vspace{1cm} {\Large #1} \nopagebreak[4]

  \noindent\rule{1cm}{0.4pt} \nopagebreak[1]}
\newcommand\peanoHeadingLarge[1]{ \vspace{2cm} {\Large \textbf{\uppercase{#1}} \nopagebreak[4]}

  \noindent\rule{2cm}{0.4pt} \nopagebreak[1]}
\newcommand\eendnoteHeading[1]{ {\Large #1} \nopagebreak[1]}


% Define commands for document structure
%
% Command for the end of each page of Peano's original book
\newcommand{\peanoPage}[1]{\vspace{1ex}

  \columnratio{0.475, 0.05, 0.475} \begin{paracol}{3} \centering \hdashrule{\columnwidth}{0.1mm}{0.1mm 1mm} \s #1 \s \hdashrule{\columnwidth}{0.1mm}{0.1mm 1mm} \end{paracol}

\vspace{1ex}}
%
% FANCY LaTeX!  Create translation environment and commands \LAT and \ENG that only work inside it.
\newenvironment{translateTwoCol}
               { % when beginning the environment
                 \columnratio{0.5, 0.5} \begin{paracol}{2}
                 \newcommand{\LAT}{\switchcolumn[0]*}
                 \newcommand{\ENG}{\switchcolumn[1]}
               }
               { % when ending the environment
                 \let\ENG\undefined
                 \let\LAT\undefined
                 \end{paracol}
               }


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


% Disable page numbering.
%   With tight margins, the numbers were in weird locations.
%   Probably want to re-enable this later, if we can.
\pagenumbering{gobble}


% Title Page
%
{ \centering

  \Large Giuseppe Peano's

  \normalsize Classic Mathematical Text

  \Huge Arithmetices principia, nova methodo exposita

  \Large OR

  \Huge ``The Principles of Arithmetic, Presented by a New Method''

  \normalsize presented

  \Large in the original Latin

  \normalsize AND

  \Large in parallel English Translation

  \vspace{1cm}

  \normalsize Original Translation By:

  \Large Vincent Verheyen

  \vspace{1cm}

  \normalsize Contributions By:

  % ADD YOUR NAME HERE
  \normalsize Michael Nahas

  \vfill

  % I also wanted to include the git revision, but it either required a Makefile
  % or it required enabling LaTeX to run shell commands, which requires an extra
  % flag on the command-line.  I thought that would be too confusing.
  
  \normalsize This document was compiled \today.

  \normalsize This document is licensed under Creative Commons Attribution-ShareAlike 4.0

  \normalsize This document is hosted at \url{https://github.com/mdnahas/Peano_Book}

} % end of centering, textsize changes
\newpage  


% About the translation/book page
%
% Center text
\columnratio{0.15, 0.7, 0.15}
\begin{paracol}{3}
  \switchcolumn[1] % use center column

  { \Large About the Translation}

   \vspace{0.25cm}
  
%  \everypar{\hangindent0.5cm}
   Below is Giuseppe Peano's \emph{Arithmetices principia} as first published\footnote{H. Kennedy, \emph{Peano. Life and Works of Giuseppe Peano}, San Francisco: Peremptory Publications, 2002, p. 41.}, i.e. as ``\emph{Arithmetices principia, nova methodo exposita}"\footnote{G. Peano, \emph{Arithmetices principia, nova methodo exposita}, Bocca, Torino, 1889.} , which appeared translated to English\footnote{These English translations listed are the only ones (to my knowledge) and all the English translations listed in: \\
   I. Grattan-Guiness (ed.), \emph{Landmark Writings in Western Mathematics 1640-1940}, Amsterdam: Elsevier, 2005, p. 614.} in 1967 as ``\emph{The principles of arithmetic, presented by a new method}"\footnote{G. Peano, (1889), ``The principles of arithmetic presented by a new method" in: J. van Heijenoort (ed.), \emph{From Frege to G\"odel. A source book in mathematical logic. 1879-1931}, Cambridge: Harvard University Press, 1967, p. 83-97.}, as well as in 1973\footnote{G. Peano, \emph{Selected works of Giuseppe Peano}, H. Kennedy (ed.), London: George Allen \& Unwin, 1973, p. 101-134.}.
   This present document\footnote{Written by Vincent Verheyen. Last updated on 17/8/2015. I encourage you to use your reason for good. If you want my support, please contact me via \url{http://vincentverheyen.com/contact}. It is possible to contribute to the flourishing of knowledge, even when you have an intelligence like mine. Thank you and good luck studying. \\
   I would like to thank Mauro Allegranzo and acknowledge his support of this work and his various comments during its creation.} is the only (to my knowledge) side-by-side Latin-English translation of the Latin original.
   The mathematical notation (in the English, right, column) got updated to currently canonically-used or easy-to-decrypt symbols in the international and/or English mathematical community; which is also a feature currently unseen in any reprint.

   \vspace{1cm}
   
   \fn{Red text} is mathematical commentary.

   \na{Gray text} is irrelavant for modern mathematical notation.

   Dashed lines (\hdashrule{2cm}{0.1mm}{0.1mm 1mm}) indicate pages in the original treatise.

   \vspace{1cm}
\end{paracol}
\newpage  



\peanoPage{I} % page-number I

\vspace{1em}

\begin{translateTwoCol}
\centering
Arithmetices principia
\ENG
The principles of arithmetic
\LAT
Nova methodo exposita
\ENG
Presented by a new method
\LAT
a
\ENG
by
\LAT
Ioseph Peano
\vspace{1em}
\ENG
Giuseppe Peano
\vspace{1em}
\LAT
in R. Academia militari professore
\ENG
\na{professor at the Royal Military Academy}
\LAT
Analysin infinitorum in R. Taurinensi Athen\ae o docente.
\ENG
\na{teaching Analysis of the infinite at the Royal Turin Athenaeum.}
\LAT
\vspace{1em}
Labor et honor
\vspace{1em}
\ENG 
\vspace{1em}
\na{Work and honor}
\vspace{1em}
\LAT
Augustae Taurinorum
\ENG
\na{At Turin}
\LAT
Ediderunt Fratres Bocca
\ENG
\na{Published by Libreria Bocca}
\LAT
Regis bibliopolae
\vspace{1em}
\ENG
...
\vspace{1em}
\end{translateTwoCol}

\columnratio{0.25, 0.25, 0.25, 0.25}
\begin{paracol}{4}
\centering
Romae \s Florentiae 
\s 
\na{At Rome} \s \na{At Florence}
\s*
Via del Corse, 216-217. \s Via Oerretani, 8.
\s
\na{Via del Corso, 216-217.} \s \na{Via Oerretani, 8.}
\end{paracol}

\begin{translateTwoCol}
\centering
1889
\ENG
\na{1889}
\end{translateTwoCol}

\peanoPage{II} % page-number II

\vspace{1em}
\begin{translateTwoCol}
\centering
Iuribus servatis
\ENG
\na{Respecting rights}
\LAT
\vspace{1em}
Augustae Taurinorum - Typis Vincentii Bona.
\ENG
\vspace{1em}
\na{At Turin, printed by Vincent Bona}
\end{translateTwoCol}

\pagebreak

\vspace{1em}
\peanoPage{III} % page-number III

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{chapter}{Preface}
\peanoHeadingLarge{Praefatio}
\ENG
\peanoHeadingLarge{Preface}
\end{translateTwoCol}

\begin{translateTwoCol}
Quaestiones, quae ad mathematicae fundamenta pertinent, etsi hisce temporibus a multis tractatae, satisfacienti solutione et adhuc carent. Hic difficultas maxime ex sermonis ambiguitate oritur.
\ENG Questions pertaining to the foundations of mathematics, although treated by many these days, still lack a satisfactory solution. The difficulty arises principally from the ambiguity of ordinary language.
\LAT
Quare summi interest verba ipsa, quibus utimur attente perpendere. Hoc examen mihi proposui, atque mei studii resultatus, et arithmeticae applicationes in hoc scripto expono.
\ENG For this reason it is of the greatest concern to consider attentively the words we use. I resolved to do this, and am presenting in this paper the results of my study with applications to arithmetic.
\LAT
Ideas omnes quae in arithmeticae principiis occurrunt, signis indicavi, ita ut quaelibet propositio his tantum signis enuncietur.
\ENG I have indicated by signs all the idea which occur in the fundamentals of arithmetic, so that every proposition is stated with just these signs.
\LAT
Signa aut ad logicam pertinent, aut proprie ad arithmeticam. Logicae signa quae hic occurrunt, sunt numero ad decem, quamvis non omnia necessaria. Horum signorum usus et proprietas nonnullae in priore parte communi sermone explicantur. Ipsorum theoriam fusius hic exponere nolui. Arithmeticae signa, ubi occurrunt, explicantur.
\ENG The signs pertain either to logic or to arithmetic. The signs of logic that occur here are about ten in number, although not all are necessary. The use of these signs and several of their properties are explained in ordinary language in the first part. I did not wish to present their theory more fully here. The signs of arithmetic are explained as they occur.
\LAT
His notationibus quaelibet propositio formam assumit atque praecisionem, qua in algebra aequationes gaudent, et a propositionibus ita scriptis aliea deducuntur, idque processis qui aequationum resolutioni assimilantur. Hoc caput totius scripti.
\ENG With this notation every proposition assumes the form and precision equations enjoy in algebra, and from propositions so written others may be deduced, by a process which resembles the solution of algebraic equations. That is the chief reason for writing this paper.
\LAT
Sique, confectis signis quibus arithmeticae propositiones scribere possim, in earum tractatione usus sum methodo, quam quia et in aliis studiis sequenda foret, breviter exponam.
\ENG Having made up the signs with which I can write arithmetical propositions, in treating them I have used a method which, because it is to be followed in later studies, I shall present briefly.
\LAT
Ex arithmeticae signis quae caeteris, una cum logicae signis exprimere licet, ideas significant quas definire possumus. Ita omnia definivi signa, si quatuor excipias, quae in explicationibus \S 1 continentur. Si, ut puto, haec ulterius reduci nequeunt, ideas ipsis expressas, ideis quae prius notae supponuntur, definire non licet.
\ENG Those arithmetical signs which may be expressed by using others along with signs of logic represent the ideas we can define. Thus I have defined every sign, if you except the four which are contained in the explanations of \S 1. If, as I believe, these cannot be reduced further, then the ideas expressed by them may not be defined by ideas already supposed to be known.
\end{translateTwoCol}

\peanoPage{IV} % page-number IV

\begin{translateTwoCol}
Propositiones, quae logicae operationibus a caeteris deducuntur, sunt \emph{theoremata}; quae vero non, \emph{axiomata} vocavi. Axiomata hic sunt novem (\S 1), et signorum, quae definitione carent, proprietates fundamentales exprimunt.
\ENG
Propositions which are deduced from others by the operations of logic are theorems; those for which this is not true I have called axioms. There are nine axioms here (\S 1), and they express fundamental properties of the undefined signs.
\LAT
In \S 1-6 numerorum proprietates communes demonstravi; brevitatis causa, demonstrationes praecedentibus similes omisi; demonstrationum communem formam immutare oportet ut logicae signis exprimantur; haec transformatio interdum difficilior est, tamen inde demonstrationis natura clarissime patet.
\ENG
In \S 1-6 I have proved the ordinary properties of numbers; for the sake of brevity, I have omitted proofs which are similar to preceding ones. The ordinary form of proofs has had to be altered in order that they may be expressed with the signs of logic. This transformation is sometimes rather difficult but the nature of the proof then becomes quite clear. In the following sections I have treated various things so that the power of the method is better seen.
\LAT
In sequentibus \S \ varia tractavi, ut huius methodi potentia magis videatur.
\ENG
In the following sections I have treated various things so that the power of the method is better seen.
\LAT
In \S 7 nonnulla theoremata, quae ad numerorum theoriam pertinent, continentur. In \S 8 et 9 rationalium et irrationalium definitiones inveniuntur.
\ENG
In \S 7 are several theorems pertaining to the theory of numbers. In \S 8 and 9 are found the definitions of rationals and irrationals.
\LAT
Denique, in \S 10, theoremata exposui nonnulla, quae nova esse puto, ad entium theoriam pertinentia, quae $\text{cl.}^{\text{mus}}$ Cantor \emph{Punktmenge (ensemble de points)} vocavit.
\ENG
Finally, in \S 10 I have given several theorems, which I believe to be new, pertaining to the theory of those entities which Professor Cantor has called \emph{Punktmenge (ensemble de points)}.
\LAT
In hoc scripto aliorum studiis usus sum. Logicae notationes et propositiones quae in num. II, III et IV continentur, si nonnullas excipias, ad multorum opera, inter quae Boole praecipue, referenda sunt.\endnote{\textbf{Giuseppe Peano's footnote (original):} \\
Boole: \\
\quad \emph{The mathematical analysis of logic ...}, Cambridge, 1847. \\
\quad \emph{The calculus of logic}, Camb. and Dublin Math. Journal, 1848. \\
\quad \emph{An investigation of the laws of thought ...}, London, 1854. \\
E. Schr\"oder: \\
\quad \emph{Der Operationskreis des Logikkalculus}, Leipzig, 1877.

Ipse iam nonnulla quae ad logicam pertinent tractavit in praecedenti opera. \\

\quad \emph{Lehrbuch der Arithmetik und Algebra ...}, Leipzig, 1873.

Boole e Schr\"oder theorias brevissime exposui in meo libro \emph{Calcolo geometrico ...}, Torino, 1888. \\
Vide: \\
\quad C. S. Pierce, \emph{On the Algebra of logic}; American Journal, III, 15; VII, 180. \\
\quad Jevons, \emph{The principles of science}, London, 1883. \\
\quad Mc.Coll., \emph{The calculus of equivalent statements}, Proceedings of the London Math. Society, 1878, Vol. IX, 9. Vol X, 16.}
\ENG
In this paper I have used the research of others. The notations and propositions of logic which are contained in numbers II, III, and IV, with some exceptions, represent the work of many, among them Boole especially.\endnote{\textbf{Giuseppe Peano's footnote (translated):} \\
Boole: \\
\quad \emph{The mathematical analysis of logic ... (Cambridge, 1847.)} \\
\quad `The calculus of logic,' \emph{Camb. and Dublin Math. J.}, 3 (1848), 193-98. \\
\quad \emph{An investigation of the laws of thought}  ... (London, 1854). \\
E. Schr\"oder: \\
\quad \emph{Der Operationskreis des Logikkalculus} (Leipzig, 1877).

He had already treated several matters pertaining to logic in a preceding work. \\

\quad \emph{Lehrbuch der Arithmetik und Algebra} ... (Leipzig, 1873).

I gave a very brief presentation of the theories of Boole and Schr\"oder in my book \emph{Calcolo geometric} etc. (Torino, 1888). \\
Cf: \\

\quad C. S. Pierce, `On the Algebra of logic,' \emph{American J. Math.}, 3 (1880), 15-57; 7 (1885), 180-202. \\
\quad Jevons, \emph{The principles of science} (London, 1883). \\
\quad Mc.Coll., `The calculus of equivalent statements,' \emph{Proc. London Math. Soc.}, 9 (1878), 9-20; 10 (1878), 16-28.}
\end{translateTwoCol}

\peanoPage{V} % page-number V

\begin{translateTwoCol}
Signum $\epsilon$, quod cum signo $\C$ confundere non licet, inversionis in logica applicationes, et paucas alias institui conventiones, ut ad exprimendam quamlibet propositionem pervenirem.
\ENG
The sign $\in$, which must not be confused with the sign $\subset$, applications of the inverse in logic, and a few other conventions, I have adopted so that I could express any proposition whatever.
\LAT
In arithmeticae demonstrationibus usus sum libro: H. Grassmann, \emph{Lehrbuch der Arithmetik}, Berlin 1861.
\ENG
In the proofs of arithmetic I used the book H. Grassmann, \emph{Lehrbuch der Arithmetik} (Berlin 1861).
\LAT
Utilius quoque mihi fuit recens scriptum: R. Dedekind, \emph{Was sind und was sollen die Zahlen}; Braunschweig, 1888, in quo quaestiones, quae ad numerorum fundamenta pertinent, acute examinantur.
\ENG
Also quite useful to me was the recent work by R. Dedekind, Was sind und was sollen die Zahlen (Braunschweig, 1888), in which questions pertaining to the foundations of numbers are acutely examined.
\LAT
Hic meus libellus ut novae methodi specimen habendus est. Hisce notationibus innumeras alias propositiones, ut quae ad rationales et irrationales pertinent, enunciare et demonstrare possumus. Sed, ut aliae theoriae tractentur, nova signa, quae nova indicant entia, instituere necesse est. Puto vero his tantum logicae signis propositiones cuiuslibet scientiae exprimi posse, dummodo adiungantur signa quae entia huius scientiae representant.
\ENG
My booklet should be taken as a sample of this new method. With these notations we can state and prove innumerable other propositions, such as those which pertain to rationals and irrationals. But in order to treat other theories, it is necessary to adopt new signs to indicate new entities. I believe, however, that with only these signs of logic the propositions of any science can be expressed, so long as the signs which represent the entities of the science are added.
\end{translateTwoCol}

\peanoPage{VI} % page-number VI

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{chapter}{Table of signs}
\peanoHeadingMedium{Signorum tabula}
\ENG
\peanoHeadingMedium{Table of signs}
\end{translateTwoCol}

\begin{translateTwoCol}
\centering
\textbf{Logicam signa}
\ENG
\textbf{Signs of logic}
\end{translateTwoCol}

\columnratio{0.05, 0.35, 0.1, 0.05, 0.35, 0.1}
\begin{paracol}{6}
\raggedright
\setlength{\columnsep}{1em}
Signum \s Significatio \s Pag.
\s Sign \s Meaning \s Pag.
\end{paracol}

\columnratio{0.06, 0.34, 0.1, 0.06, 0.34, 0.1}
\begin{paracol}{6}
\raggedright
\setlength{\columnsep}{1em}
$P$ \s propositio \s VII
\s \scriptsize{proposition} \normalsize \s proposition \s VII
\s* $K$ \s classis \s X
\s \scriptsize{CLASS} \normalsize \s class \s X
\s* $\cap$ \s et \s VII, X
\s $\wedge$ \s and \s VII, X
\s* $\cup$ \s vel \s VIII, X, XI
\s $\vee$ \s or \s VIII, X, XI
\s* $-$ \s non \s VIII, X
\s $\neg$ \s not \s VIII, X
\s* $\abs$ \s absurdum \emph{aut} nihil \s VIII, XI
\s $\bot$ \\ $\varnothing$ \s false \emph{or} \\ \quad nothing \s VIII, XI
\s* $\C$ \s deducitur \emph{aut} continetur \s VIII, XI
\s $\rightarrow$ \\ $\subset$ \s one deduces \emph{or} \\ \quad is contained in \s VIII, XI
\s* = \s est aequalis \s VIII
\s = \s equals \s VIII
\s* $\epsilon$ \s est \s X
\s $\in$ \\ is a \s is (an element of) \emph{or} \\ is (a) \s X
\s* $[ \ ]$ \s inversionis signum \s XI
\s $\{ x | \ldots \}$ \s sign of the inverse \s XI
\s* \rotatebox[origin=c]{180}{$\epsilon$} \s qui \emph{vel} $[ \epsilon ]$ \s XII
\s $\rightarrow$ \\ isn't \s such that \emph{or} \\ \quad isn't
\s XII
\s* Th \s Theorema \s XVI
\s \scriptsize{theorem} \s \normalsize Theorem \s XVI
\s* Hp \s Hypothesis \s
\s \scriptsize{hypothesis} \s \normalsize Hypothesis \s
\s* Ts \s Thesis \s
\s \scriptsize{thesis} \s \normalsize Thesis \s
\s* L \s Logica \s
\s \scriptsize{logic} \s \normalsize Logic \s
\end{paracol}

\vspace{1em}
\begin{translateTwoCol}
\centering
\textbf{Arithmeticae signa}
\ENG
\textbf{Signs of arithmetic}
\end{translateTwoCol}

\columnratio{0.05, 0.35, 0.1, 0.05, 0.35, 0.1}
\begin{paracol}{6}
\raggedright
\setlength{\columnsep}{1em}
Signum \s Significatio \s Pag.
\s Sign \s Meaning \s Pag.
\end{paracol}

\begin{translateTwoCol}
\raggedright
Signa 1, 2, ..., =, $>$, $<$, +, -, $\times$ vulgarem habent significationem. Divisionis signum est /.
\ENG
The signs 1, 2, ..., =, $>$, $<$, +, -, $*$ have their usual meaning. The sign of division is /.
\end{translateTwoCol}

\columnratio{0.06, 0.34, 0.1, 0.06, 0.34, 0.1}
\begin{paracol}{6}
\raggedright
\setlength{\columnsep}{1em}
$N$ \s numerus integer positivus \s 1
\s $\mathbb{N}$ \s positive integers \s 1
\s* $R$ \s num. rationalis positivus \s 12
\s $\mathbb{R}$ \s postive rational numbers \s 12
\s* $Q$ \s quantitas, \emph{sive} numerus realis positivus \s 16
\s $\mathbb{Q}$ \s quantity \emph{or} postive real numbers \s 16
\s* $\Np$ \s numerus primus \s 9
\s $\mathbb{P}$ \s prime number \s 9
\s* M \s maximus \s 6
\s M$^{\text{aximum}}$ \s maximum \s 6
\s* $\mini$ \s minimus \s 6
\s m$_{\text{inimum}}$ \s minimum \s 6
\s* T \s terminus, \emph{vel} limes summus \s 15
\s Gr$^{\text{eatest bound}}$ \s terminus \emph{or} greatest bound \s 15
\s* D \s dividit \s 9
\s \scriptsize{divides} \s \normalsize divides
\s 9
\s* \reflectbox{D} \s est multiplex \s 9
\s \scriptsize{is divisible} \s \normalsize is divisible \s 9
\s* $\pi$ \s est primus cum \s 6
\s \scriptsize{is $\mathbb{P}$rime with} \s \normalsize is prime with \s 6
\end{paracol}

\vspace{1em}
\begin{translateTwoCol}
\centering
\textbf{Signa composita}
\ENG
\textbf{Composite signs}
\end{translateTwoCol}

\columnratio{0.06, 0.34, 0.1, 0.06, 0.34, 0.1}
\begin{paracol}{6}
\raggedright
\setlength{\columnsep}{1em}
$-<$ \s non est minor \s
\s $\nless$ \s is not less than \s
\s* $=\cup>$ \s est aequalis aut maior \s
\s $\geq$ \s is equal to or greater than \s
\s* \rotatebox[origin=c]{180}{$\epsilon$} $D$ \s divisor \s
\s \scriptsize{is a divisor} \s \normalsize{is a divisor} \s
\s* $M$ \rotatebox[origin=c]{180}{$\epsilon$} $D$ \s maximus divisor \s
\s \scriptsize{is the greatest divisor} \s \normalsize is the greatest divisor \s
\end{paracol}

\peanoPage{VII} % page-number VII

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{chapter}{Notations of logic}
\peanoHeadingMedium{Logicae notationes.}
\ENG
\peanoHeadingMedium{Notations of logic.}
\end{translateTwoCol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{section}{I. Punctuation}
\peanoHeadingSmall{I. De punctuatione.}
\ENG
\peanoHeadingSmall{I. Punctuation.}
\end{translateTwoCol}

\begin{translateTwoCol}
Litteris $a,b,... x,y,... x', y',...$ entia indicamus indeterminata quaecumque. Entia vero determinata signis, sive litteris $P, K, N,...$ indicamus.
\ENG
By the letters $a,b,... x,y,... x', y',...$ we indicate any indeterminate entities. Determinate entities are, however, indicated by the signs, or rather by the letters, $P, K, N,...$
\LAT
Signa plerumque in eadem linea scribemus. Ut ordo pateat quo ea coniungere oporteat, \emph{parenthesibus} ut in algebra, sive \emph{punctis} $. : \pppNoSpace :\hspace{0.1em}:$ etc. utimur.
\ENG
\na{Generally we write signs on the same line. So that it will be clear how they are to be joined, we use \emph{parentheses}, as in algebra, or rather \emph{points} $. : \pppNoSpace :\hspace{0.1em}:$ etc.}
\LAT
Ut formula punctis divisa, intelligatur, primum signa quae nullo puncto seperantur colligenda sunt, postea quae uno puncto, deinde quae duobus punctis, etc.
\ENG
\na{So that a formula divided by points may be understood, first the signs which are not seperated by points are taken together, then those seperated by one point, then those by two points, etc.}
\LAT
Ex. g. sint $a,b,c,...$ signa quaecumque. Tunc $ab$ $.$ $cd$ significat $(ab)(cd)$; et $ab$ $.$ $cd:ef$ $.$ $gh$ $\pppNoSpace$ $k$ significat $(((ab)(cd))((ef)(gh)))k$.
\ENG
\na{For example, let $a,b,c,...$ be any signs. Then $(ab)(cd)$ means $(ab)(cd)$; and $\{[(ab)(cd)][(ef)(gh)]\}k$ means $\{[(ab)(cd)][(ef)(gh)]\}k$.}
\LAT Punctuationis signa omittere licet si formulae quae diversa punctuatione existerent eundem habeant sensum; vel si una tantum formula, et ipsa quam scribere volumus, sensum habeat.
\ENG \na{The signs of punctuation may be omitted if formulas having different punctuation have the same meaning, or if just one formula, that being the one we wish to write, has meaning.}
\LAT Ut ambiguitatis periculum absit, aritmeticae operationum signis $. :$ nunquam utimur.
\ENG \na{To avoid the danger of ambiguity, we never use $. :$ as signs of arithmetical operations.}
\LAT Parenthesum figura una est $( \ )$; si in eadem formula, parentheses et puncta occurant, primum quae parenthesibus continentur, colligantur.
\ENG \na{The figure of parentheses is one $( \ )$; if parentheses and points occur in the same formula, whatever is contained in parentheses is to be gathered first.}
\end{translateTwoCol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{section}{II. Propositions}
\peanoHeadingSmall{II. De propositionibus.}
\ENG
\peanoHeadingSmall{II. Propositions.}
\end{translateTwoCol}

\begin{translateTwoCol}
Signo $P$ significatur \emph{propositio}.
\ENG
The sign $P$ means \emph{proposition}.
\LAT
Signum $\cap$ legitur \emph{et}. Sint $a, b$ propositiones; tunc $a \cap b$ est simultanea affirmatio propositionum $a, b$. Brevitatis causa, loco $a \cap b$ vulgo scribemus $a$ $b$.
\ENG
The sign $\wedge$ is read \emph{and}. Let $a, b$, be propositions; then $a \wedge b$ is the simultaneous affirmation of the propositions $a, b$. For the sake of brevity, instead of $a \wedge b$, we ordinarily write $a,b$.
\end{translateTwoCol}

\peanoPage{VIII} % page-number VIII

\begin{translateTwoCol}
Signum $-$ legitur \emph{non}. Sit $a$ quaedam $P$; tunc $-a$ est negatio propositionis $a$.
\ENG The sign $\neg$ is read \emph{not}. Let $a$ be a $P$; then $\neg a$ is the negation of the proposition $a$.
\LAT Signo $\cup$ legitur \emph{vel}. Sint $a,b$ propositiones; tunc $a \cup b$ idem est ac $-:-a$ $.$ $-b$.
\ENG The sign $\vee$ is read \emph{or}. Let $a,b$ be propositions; then $a \vee b$ is the same as $\neg[(\neg a) \wedge (\neg b)]$.
\LAT \text{[}Signo $V$ significatur \emph{verum}, sive \emph{identitas}; sed hoc signo numquam utimur\text{]}.  % TODO: The math V here should be upright, not slanted
\ENG \na{The sign $\top$  means \emph{true}, or \emph{identity}, but we never use this sign.}
\LAT Signum  $\abs$ significat \emph{falsum}, sive \emph{absurdum}.
\ENG The sign $\bot$ means \emph{false}, or \emph{absurd}.
\LAT \text{[}Signum C significat \emph{est consequentia}; ita $b$ C $a$ legitur $b$ \emph{est consequentia propositionis} $a$. Sed hoc signo nunquam utimur\text{]}.
\ENG \na{The sign $\leftarrow$ means \emph{is a consequence of}. Thus $b \leftarrow a$ is read $b$ \emph{is a consequence of the proposition} $a$. But we never use this sign.}
\LAT Signum $\C$ significat \emph{deducitur}; ita $a$ $\C$ $b$ significat quod $b$ C $a$. Si propositiones $a$, $b$ entia indeterminata continent $x$, $y$,..., scilicet sunt inter ipsa entia conditiones, tunc $a$ $\C_{x,y,...}$ $b$ significat: quaecumque sunt $x,y,...,$ a propositione $a$ deducitur $b$. Si vero ambiguitatis periculum absit, loco $\C_{x,y,...}$, scribemus solum $\C$.
\ENG The sign $\rightarrow$ means \emph{one deduces}; thus $a \rightarrow b$ means the same as $b \leftarrow a$. If the propositions $a,b$ contain the indeterminate quantities $x,y,...$, that is, express conditions on these objects, then $a \xrightarrow[x,y,...]{} b$ means: whatever the $x$, $y$,..., from propositions $a$ one deduces $b$. If indeed there is no danger of ambiguity, instead of $\xrightarrow[x,y,...]{}$, we write only $\rightarrow$.
\LAT Signum = significat \emph{est aequalis}. Sint $a,b$ propositiones; tunc $a=b$ idem significat quod $a$ $\C$ $b$ $.$ $b$ $\C$ $a$; propositio $a=_{x,y,...} b$ idem significat quod $a$ $ \C_{x,y,...} b$ $.$ $b$ $\C_{x,y,...} a$.
\ENG The sign = means \emph{equals}. Let $a,b$ be propositions; then $a=b$ means the same as $(a \rightarrow b) \wedge (b \rightarrow a)$; proposition $a\underset{x,y,...}=b$ means the same as ($a \xrightarrow[\text{x,y,...}]{} b) \wedge (b \xrightarrow[x,y,...]{} a)$.
\end{translateTwoCol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{section}{III. Propositions of logic}
\peanoHeadingSmall{III. Logicae propositiones.}
\ENG
\peanoHeadingSmall{III. Propositions of logic.}
\end{translateTwoCol}

\begin{translateTwoCol}
Sint $a,b,c,...$ propositiones. Tunc erit:
\ENG
Let $a,b,c,...$ be propositions. We have:
\LAT
1. \hspace{0.6cm} $a$ $\C$ $a$
\ENG
1. \hspace{0.6cm} $a \rightarrow a$
\LAT
2. \hspace{0.6cm} $a$ $\C$ $b$ . $b$ $\C$ $c$ : $\C$ : $a$ $\C$ $c$
\ENG
2. \hspace{0.6cm} $[(a \rightarrow b) \wedge (b \rightarrow c)] \rightarrow (a \rightarrow c)$
\LAT
3. \hspace{0.6cm} $a$ = $b$ . = : $a$ $\C$ $b$ . $b$ $\C$ $a$.
\ENG
3. \hspace{0.6cm} $(a=b)=[(a \rightarrow b) \wedge (b \rightarrow a)]$
\LAT
4. \hspace{0.6cm} $a$ = $a$
\ENG
4. \hspace{0.6cm} $a=a$
\LAT
5. \hspace{0.6cm} $a$ = $b$ . = . $b$ = $a$
\ENG
5. \hspace{0.6cm} $(a=b)=(b=a)$
\LAT
6. \hspace{0.6cm} $a$ = $b$ . $b$ $\C$ $c$ : $\C$ . $a$ $\C$ $c$
\ENG
6. \hspace{0.6cm} $[(a=b) \wedge (b \rightarrow c)] \rightarrow (a \rightarrow c)$
\LAT
7. \hspace{0.6cm} $a$ $\C$ $b$ . $b$ = $c$ : $\C$ . $a$ $\C$ $c$
\ENG
7. \hspace{0.6cm} $[(a \rightarrow b) \wedge (b=c)] \rightarrow (a \rightarrow c)$
\LAT
8. \hspace{0.6cm} $a$ = $b$ . $b$ = $c$ : $\C$ . $a$ = $c$
\ENG
8. \hspace{0.6cm} $[(a=b) \wedge (b=c) ] \rightarrow (a=c)$
\LAT
9. \hspace{0.6cm} $a$ = $b$ . $\C$ . $a$ $\C$ $b$
\ENG
9. \hspace{0.6cm} $(a=b) \rightarrow (a \rightarrow b)$
\LAT
10. \hspace{0.67cm} $a$ = $b$ . $\C$ . $b$ $\C$ $a$
\ENG
10. \hspace{0.67cm} $(a=b) \rightarrow (b \rightarrow a)$
\LAT
\hfill
\ENG
\hfill
\LAT
11. \hspace{0.67cm} $ab$ $\C$ $a$
\ENG
11. \hspace{0.67cm} $(a \wedge b) \rightarrow a$
\LAT
12. \hspace{0.67cm} $ab$ = $ba$
\ENG
12. \hspace{0.67cm} $(a \wedge b)=(b \wedge a)$
\LAT
13. \hspace{0.67cm} $a$\thinspace $(bc)$ = $(ab)$\thinspace $c$ = $abc$
\ENG
13. \hspace{0.67cm} $(a \wedge (b \wedge c))=((a \wedge b) \wedge c)=(a \wedge b \wedge c)$
\end{translateTwoCol}

\peanoPage{IX} % page-number IX

\begin{translateTwoCol}
14. \hspace{0.67cm} $aa$ = $a$
\ENG
14. \hspace{0.67cm} $(a \wedge a)=a$
\LAT
15. \hspace{0.67cm} $a$ = $b$ . $\C$ . $ac$ = $bc$
\ENG
15. \hspace{0.67cm} $(a=b) \rightarrow [(a \wedge c)=(b \wedge c)]$
\LAT
16. \hspace{0.67cm} $a$ $\C$ $b$ . $\C$ . $ac$ $\C$ $bc$
\ENG
16. \hspace{0.67cm} $(a \rightarrow b) \rightarrow [(a \wedge c) \rightarrow (b \wedge c)]$
\LAT
17. \hspace{0.67cm} $a$ $\C$ $b$ . $c$ $\C$ $d$ : $\C$ . $ac$ $\C$ $bd$
\ENG
17. \hspace{0.67cm} $[(a \rightarrow b) \wedge (c \rightarrow d)] \rightarrow [(a \wedge c) \rightarrow (b \wedge d)]$
\LAT
18. \hspace{0.67cm} $a$ $\C$ $b$ . $a$ $\C$ $c$ : = . $a$ $\C$ $bc$
\ENG
18. \hspace{0.67cm} $[(a \rightarrow b) \wedge (a \rightarrow c)]=[(a \rightarrow (b \wedge c)]$
\LAT
19. \hspace{0.67cm} $a$ = $b$ . $c$ = $d$ : $\C$ . $ac$ = $bd$
\ENG
19. \hspace{0.67cm} $[(a=b) \wedge (c=d)] \rightarrow [(a \wedge c)=(b \wedge d)]$
\LAT
\hfill
\ENG
\hfill
\LAT
20. \hspace{0.67cm} $-(-a)$ = $a$
\ENG
20. \hspace{0.67cm} $\neg(\neg a) = a$
\LAT
21. \hspace{0.67cm} $a$ = $b$ . = . $-a$ = $-b$.
\ENG
21. \hspace{0.67cm} $(a=b)=[(\neg a)=(\neg b)]$
\LAT
22. \hspace{0.67cm} $a$ $\C$ $b$ . = . $-b$ $\C$ $-a$ =
\ENG
22. \hspace{0.67cm} $(a \rightarrow b)=[(\neg b) \rightarrow (\neg a)]$
\LAT
\hfill
\ENG
\hfill
\LAT
23. \hspace{0.67cm} $a$ $\cup$ $b$ . = $\pppNoSpace$ $-$ : $-a$ . $-b$
\ENG
23. \hspace{0.67cm} $(a \vee b) = \{[(\neg[(\neg a) \wedge (\neg b)]\} $
\LAT
24. \hspace{0.67cm} $-(ab)$ = $(-a)$ $\cup$ $(-b)$
\ENG
24. \hspace{0.67cm} $[\neg(a \wedge b)] = [(\neg a) \vee (\neg b)]$
\LAT
25. \hspace{0.67cm} $-(a$ $\cup$ $b)$ = $(-a)$  $(-b)$
\ENG
25. \hspace{0.67cm} $[\neg(a \vee b)] = [(\neg a) \vee (\neg b)]$
\LAT
26. \hspace{0.67cm} $a$ $\C$ . $a$ $\cup$ $b$
\ENG
26. \hspace{0.67cm} $a \rightarrow (a \vee b)$
\LAT
27. \hspace{0.67cm} $a$ $\cup$ $b$ = $b$ $\cup$ $a$
\ENG
27. \hspace{0.67cm} $(a \vee b) = (b \vee a)$
\LAT
28. \hspace{0.67cm} $a$ $\cup$ $(b$ $\cup$ $c)$ = $(a$ $\cup$ $b)$ $\cup$ $c$ = $a$ $\cup$ $b$ $\cup$ $c$
\ENG
28. \hspace{0.67cm} $[a \vee (b \vee c)] = [(a \vee b) \vee c] = (a \vee b \vee c)$
\LAT
29. \hspace{0.67cm} $a$ $\cup$ $a$ = $a$
\ENG
29. \hspace{0.67cm} $(a \vee a)=a$
\LAT
30. \hspace{0.67cm} $a$ $(b$ $\cup$ $c)$ = $ab$ $\cup$ $ac$
\ENG
30. \hspace{0.67cm} $[a \wedge (b \vee c)]=[(a \wedge b) \vee (a \wedge c)]$
\LAT
31. \hspace{0.67cm} $a$ = $b$ . $\C$ . $a$ $\cup$ $c$ = $b$ $\cup$ $c$
\ENG
31. \hspace{0.67cm} $(a=b) \rightarrow [(a \vee c) = (b \vee c)]$
\LAT
32. \hspace{0.67cm} $a$ $\C$ $b$ . $\C$ . $a$ $\cup$ $c$ $\C$ $b$ $\cup$ $c$
\ENG
32. \hspace{0.67cm} $(a \rightarrow b) \rightarrow [(a \vee c) \rightarrow (b \vee c)]$
\LAT
33. \hspace{0.67cm} $a$ $\C$ $b$ . $c$ $\C$ $d$ : $\C$ : $a$ $\cup$ $c$ . $\C$ . $b$ $\cup$ $d$
\ENG
33. \hspace{0.67cm} $[(a \rightarrow b) \wedge (c \rightarrow d)] \rightarrow [(a \vee c) \rightarrow (b \vee d)]$
\LAT
34. \hspace{0.67cm} $b$ $\C$ $a$ . $c$ $\C$ $a$ : = . $b$ $\cup$ $c$ $\C$ $a$
\ENG
34. \hspace{0.67cm} $[(b \rightarrow a) \wedge (c \rightarrow a)] = [(b \vee c) \rightarrow a]$
\LAT
\hfill
\ENG
\hfill
\LAT
35. \hspace{0.67cm} $a-a$ = $\abs$
\ENG
35. \hspace{0.67cm} $[a \wedge (\neg a)] = \bot$
\LAT
36. \hspace{0.67cm} $a$ $\abs$ = $\abs$
\ENG
36. \hspace{0.67cm} $(a \wedge \bot) = \bot$
\LAT
37. \hspace{0.67cm} $a$ $\cup$ $\abs$ = a
\ENG
37. \hspace{0.67cm} $(a \vee \bot)=a$
\LAT
38. \hspace{0.67cm} $a$ $\C$ $\abs$ . = . $a$ = $\abs$
\ENG
38. \hspace{0.67cm} $(a \rightarrow \bot) = (a = \bot)$
\LAT
39. \hspace{0.67cm} $a$ $\C$ $b$ . = . $a-b$ = $\abs$
\ENG
39. \hspace{0.67cm} $(a \rightarrow b) = [(a \wedge \neg b) = \bot]$
\LAT
40. \hspace{0.67cm} $\abs$ $\C$ $a$
\ENG
40. \hspace{0.67cm} $\bot \rightarrow a$
\LAT
41. \hspace{0.67cm} $a$ $\cup$ $b$ = $\abs$ . = : $a$ = $\abs$ . $b$ = $\abs$
\ENG
41. \hspace{0.67cm} $[(a \vee b)= \bot]=[(a= \bot) \wedge (b= \bot)]$
\LAT
\hfill
\ENG
\hfill
\LAT
42. \hspace{0.67cm} $a$ $\C$ . $b$ $\C$ $c$ : = : $ab$ $\C$ $c$
\ENG
42. \hspace{0.67cm} $[a \rightarrow (b \rightarrow c)] = [(a \wedge b) \rightarrow c]$
\LAT
43. \hspace{0.67cm} $a$ $\C$ . $b$ = $c$ : = . $ab$ = $ac$
\ENG
43. \hspace{0.67cm} $[a \rightarrow (b=c)]=[(a \wedge b)=(a \wedge c)]$
\end{translateTwoCol}

\peanoPage{X} % page-number X

\begin{translateTwoCol}
\quad Sit $\alpha$ quoddam relationis signum (ex. gr. $=$, $\C$), ita ut $a$ $\alpha$ $b$ sit quaedam propositio. Tunc loco $-$ . $a$ $\alpha$ $b$ scribemus $a$ $-$ $\alpha$ $b$; scilicet:
\ENG
\quad Let $\alpha$ be the sign of some relation (eg., $=, \rightarrow$) so that $a$ $\alpha$ $b$ is a proposition. Then instead of $\neg(a$ $\alpha$ $b)$, we write $a$ $\not\alpha$ $b$. Thus:
\LAT
\hspace{1.06cm} $a$ $-$ = $b$ . = : $-$ . $a$ = $b$
\ENG
\hspace{1.06cm} $(a \not= b) = \neg (a=b)$
\LAT
\hspace{1.06cm} $a$ $-$ $\C$ $b$ . = : $-$ . $a$ $\C$ $b$
\ENG
\hspace{1.06cm} $ (a \not\rightarrow b) = \neg (a \rightarrow b)$
\LAT
\quad Ita signum $-$ $=$ significat {\emph{non est aequalis}}. Si propositio $a$ indeterminatum continet $x$, $a$ $-=$\scalebox{0.7}{$x$}\thinspace $\abs$ significat: sunt $x$ quae conditioni $a$ satisfaciunt. Signum $-$ $\C$ significat {\emph{non deducitur}}.
\ENG
Thus the sign $\not=$ means {\emph{is not equal to}}. If the proposition $a$ contains the indeterminate $x, a\underset{\text{x}}\neq \bot$ means: there is an $x$ which satisfies condition $a$. The sign $\not\rightarrow$ means {\emph{one does not deduce}}.
\LAT
\quad Similter, si $\alpha$ et $\beta$ sunt relationis signa, loco $a$ $\alpha$ $b$, et $a$ $\alpha$ $b$ . $\cup$ . $a$ $\beta$ $b$ scribere possumus $a$ . $\alpha$ $\beta$ . $b$ et $a$ . $\alpha$ $\cup$ $\beta$ . $b$. Ita, si $a$ et $b$ sunt propositiones, formula $a$ . $\C$ $-$ = . $b$ dicit: ab $a$ {\emph{deducitur}} $b$, sed non vice versa.
\ENG
\quad Similarly, if $\alpha$ and $\beta$ are signs of relations, instead of $(a$ $\alpha$ $b) \wedge (a$ $\beta$ $b)$, and $(a$ $\alpha$ $b) \vee (a$ $\beta$ $b)$ we may write $a$ $(\alpha \wedge \beta)$ $b$ and $a$ $(\alpha \vee \beta)$ $b$. Thus, if $a$ and $b$ are propositions, the formula $a$ $(\rightarrow \wedge \not=)$ $b$ says: from $a$ {\emph{one deduces}} $b$, but not vice versa.
\LAT
\hspace{1.06cm} $a$ . $\C$ $-$ = . $b$ : = : $a$ $\C$ $b$ . $b$ $-$ $$\C$$ $a$
\ENG
\hspace{1.06cm} $[a$ $(\rightarrow \wedge \not=)$ $b] = [(a \rightarrow b) \wedge (b \not\rightarrow a)]$
\LAT
\quad Formulae:
\ENG
\quad Formulas:
\LAT
\hspace{1.06cm} $a$ $\C$ $b$ . $b$ $\C$ $c$ . $a$ $-$ $\C$ $c$ : = $\abs$
\ENG
\hspace{1.06cm} $[(a \rightarrow b) \wedge (b \rightarrow c) \wedge (a \not\rightarrow c)] = \bot$
\LAT
\hspace{1.06cm} $a$ = $b$ . $b$ = $c$ . $a$ $-$ = $c$ : = $\abs$
\ENG
\hspace{1.06cm} $[(a=b) \wedge (b=c) \wedge (a \not= c)] = \bot$
\LAT
\hspace{1.06cm} $a$ $\C$ $b$ . $b$ $\C$ $-$ = $c$ : $\C$ . $a$ $\C$ $-$ = $c$
\ENG
\hspace{1.06cm} $\{(a \rightarrow b) \wedge [b$ $(\rightarrow \wedge \not=)$ $c]\} \rightarrow [a$ $(\rightarrow \wedge \not=)$ $c]$
\LAT
\hspace{1.06cm} $a$ $\C$ $-$ = $b$ . $b$ $\C$ $c$ : $\C$ . $a$ $\C$ $-$ = $c$
\ENG
\hspace{1.06cm} $[a$ $(\rightarrow \wedge \not=)$ $b] \wedge (b \rightarrow c)\} \rightarrow [a$ $(\rightarrow \wedge \not=)$ $c]$
\LAT
\quad Sed his notationibus raro utimur.
\ENG
\quad But we shall rarely use these notations.
\end{translateTwoCol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{section}{IV. Classes}
\peanoHeadingSmall{IV. De classibus.}
\ENG
\peanoHeadingSmall{IV. Classes.}

\raggedright
\quad \fn{COMMENTARY: Peano's ``Class'' is similar to modern ``sets''. }
\end{translateTwoCol}

\begin{translateTwoCol}
\quad Signo K significatur {\emph{classis}}, sive entium aggregatio.
\ENG
\quad The sign \scriptsize{CLASS} \normalsize means {\emph{class}}, or aggregate of entities.
\LAT
\quad Signum \scalebox{0.7}[0.9]{$\in$} significat {\emph{est}}. Ita $a$ \scalebox{0.7}[0.9]{$\in$} $b$ legitur $a$ {\emph{est quoddam}} $b$; $a$ $\epsilon$ K significat $a$ {\emph{est quaedam classis}}; $a$ \scalebox{0.7}[0.9]{$\in$} P significat $a$ {\emph{est quaedam propositio}}.
\ENG
\quad The sign $\in$ means {\emph{is an element of}}. Thus $a \in b$ is read $a$ {\emph{is an element of}} $b$\na{; $A$ is a \scriptsize{CLASS} \normalsize means {\emph{$A$ is a class}}; $a$ is a \scriptsize{proposition} \normalsize means {\emph{a is a proposition}}}.
\LAT
\quad Loco $-(a$ \scalebox{0.7}[0.9]{$\in$} $b)$ scribemus $a$ $-$ \scalebox{0.7}[0.9]{$\in$} $b$; signum $-$ \scalebox{0.7}[0.9]{$\in$} significat {\emph{non est}}; scilicet:
\ENG
\quad Instead of $\neg(a \in b)$ we shall write $a \not\in b$. The sign $\not\in$ means {\emph{is not}}; thus:
\LAT
44. \hspace{0.67cm} $a$ $-$ \scalebox{0.7}[0.9]{$\in$} $b$ . = : $-$ . $a$ \scalebox{0.7}[0.9]{$\in$} $b$
\ENG
44. \hspace{0.67cm} $(a \not\in b) = [\neg(a \in b)]$
\LAT
\quad Signum $a$, $b$, $c$ \scalebox{0.7}[0.9]{$\in$} $m$ significat: $a$, $b$ et $c$ sunt $m$; scilicet:
\ENG
\quad The sign $(a, b, c) \in m$ means: $a$, $b$, and $c$ are in $m$; thus:
\LAT
45. \hspace{0.67cm} $a$, $b$, $c$ \scalebox{0.7}[0.9]{$\in$} $m$ . = : $a$ \scalebox{0.7}[0.9]{$\in$} $m$ . $b$ \scalebox{0.7}[0.9]{$\in$} $m$ . c \scalebox{0.7}[0.9]{$\in$} $m$
\ENG
45. \hspace{0.67cm} $[(a, b, c) \in m] = [(a \in m) \wedge (b \in m) \wedge (c \in m)]$
\LAT
\quad Sit $a$ classis; tunc $-a$ significatur classis indiviuis constituta quae non sunt $a$.
\ENG
\quad Let $A$ be a class. Then $\overline{A}$ means that class made up of individuals that are not in $A$. 
\LAT
46. \hspace{0.67cm} $a$ \scalebox{0.7}[0.9]{$\in$} K . $\C$ : $x$ \scalebox{0.7}[0.9]{$\in$} $-$ $a$ . = . $x$ $-$ \scalebox{0.7}[0.9]{$\in$} $a$
\ENG
46. \hspace{0.67cm} $A$ is a \scriptsize{CLASS} \normalsize $\rightarrow [(x \in \overline{A}) = (x \not\in A)]$
\LAT
\quad Sint $a$, $b$ classes; $a$ \scalebox{0.8}[0.6]{$\cap$} $b$, sive $a$ $b$, est classis individuis constituta
\ENG
\quad Let $A, B$ be classes. Then $A \cap B$\na{, or $A B$,} is the class composed of individuals
\end{translateTwoCol}

\peanoPage{XI} % page-number XI

\begin{translateTwoCol}
quae eodem tempore sunt $a$ et $b$; $a$ $\cup$ $b$ est classis individuis constituta qui sunt $a$ vel $b$.
\ENG
which are at the same time in $A$ and $B$; $A \cup B$ is the class composed of individuals which are in $A$ or $B$.
\LAT
47. \hspace{0.67cm} $a$, $b$ \scalebox{0.7}[0.9]{$\in$} K . $\C$ $\pppNoSpace$ $a$ $x$ \scalebox{0.7}[0.9]{$\in$} . $a$ $b$ : = : $x$ \scalebox{0.7}[0.9]{$\in$} $a$ . $x$ \scalebox{0.7}[0.9]{$\in$} $b$
\ENG
47. \hspace{0.67cm} $A, B$ are \scriptsize{CLASSES} \normalsize $\rightarrow \{[(x \in (A \cap B)] = [(x \in A) \wedge (x \in B)]\}$
\LAT
48. \hspace{0.67cm} $a$, $b$ \scalebox{0.7}[0.9]{$\in$} K . $\C$ $\pppNoSpace$ $a$ $\cup$ $x$ \scalebox{0.7}[0.9]{$\in$} . $a$ $\cup$ $b$ : = : $x$ \scalebox{0.7}[0.9]{$\in$} $a$ . $\cup$ . $x$ \scalebox{0.7}[0.9]{$\in$} $b$
\ENG
48. \hspace{0.67cm} $A, B$ are \scriptsize{CLASSES} \normalsize $\rightarrow \{[(x \in (A \cup B)] = [(x \in A) \vee (x \in B)]\}$
\LAT
\quad Signum $\abs$ indicat classem quae nullum continet individuum. Ita:
\ENG
\quad The sign $\varnothing$ indicates the class which contains no individuals. Thus:
\LAT
49. \hspace{0.67cm} $a$ \scalebox{0.7}[0.9]{$\in$} K . $\C$ $\pppNoSpace$ $a$ = $\abs$ : = : $x$ \scalebox{0.7}[0.9]{$\in$} $a$ . =\scalebox{0.7}{$x$}\thinspace  $\abs$
\ENG
49. \hspace{0.67cm} $A$ is a \scriptsize{CLASS} \normalsize $\rightarrow \{(A= \varnothing) = [(x \in A) \underset{\forall \text{x}}= \bot])\}$ 
\LAT
\quad [Signo V, quod classem ex omnibus individuis constitutam, de quibus quaestio est, indieat, non utimur].
\ENG
\quad \na{[We shall not use the sign \scriptsize{true} \normalsize, which indicates the class composed of all individuals being considered].}
% TODO Find a replacement for ``true'' above
\LAT
\quad Signum $\C$ significat {\emph{continetur}}. Ita $a$ $\C$ $b$ significat {\emph{classis a continetur in classi b}}.
\ENG
\quad The sign $\subset$ means {\emph{is contained}}. Thus $a \subset b$ means {\emph{the class a is contained in the class b}}.
\LAT
50. \hspace{0.67cm} $a$, $b$ \scalebox{0.7}[0.9]{$\in$} K . $\C$ $\pppNoSpace$ $a$ $\C$ $b$ : = : $x$ \scalebox{0.7}[0.9]{$\in$} $a$ . $\C$\scalebox{0.7}{$x$}\thinspace . $x$ \scalebox{0.7}[0.9]{$\in$} $b$
\ENG
50. \hspace{0.67cm} $A, B$ are \scriptsize{CLASSES} \normalsize $\rightarrow \{(A \subset B)=[(x\in A) \xrightarrow[\forall \text{x}]{} (x \in B)]\}$
\LAT
\quad [Formula $b$ \scalebox{0.8}{C} $a$ significare potest {\emph{classis b continet classem a}}; at signo  \scalebox{0.8}{C} non utiumur].
\ENG
\quad \na{[The formula $B \supset A$ could mean {\emph{the class B contains the class A}}, but we shall not use the sign $\supset$].}
\LAT
Hic signa $\abs$ et $\C$ significationem habent quae paullo a praecedenti differt; sed nulla orietur ambiguitas. Nam si de propositionibus agatur, haec signa legantur {\emph{absurdum}} et {\emph{deducitur}}; si vero de classibus, {\emph{nihil}} et {\emph{continetur}}.
\ENG
\na{The signs $(\bot / \varnothing)$ and $(\rightarrow / \subset)$ have meanings here which are slightly different from the preceding, but no ambiguity will arise, for if propositions are being considered, the signs are read \emph{absurd} $(\bot)$ and \emph{one deduces} $(\rightarrow)$, but if classes are being considered, they are read \emph{empty} $(\varnothing)$ and \emph{is contained} $(\subset)$.}
\LAT
\quad Formula $a$ = $b$, si $a$ et $b$ sint classes, significat $a$ $\C$ $b$ . $b$ $\C$ $a$. Itaque
\ENG
\raggedright
\quad The formula $A = B$, if $A$ and $B$ are classes, means $(A \subset B) \wedge (B \subset A)$. Thus
\LAT
51. \hspace{0.67cm} $a$, $b$ \scalebox{0.7}[0.9]{$\in$} K . $\C$ $\pppNoSpace$ $a$ = $b$ : = : $x$ \scalebox{0.7}[0.9]{$\in$} $a$ . =\scalebox{0.7}{$x$}\thinspace . $x$ \scalebox{0.7}[0.9]{$\in$} $b$
\ENG
51. \hspace{0.67cm} $A$, $B$ are \scriptsize{CLASSES} \normalsize $\rightarrow \{(A = B) = [(x \in A) \underset{\forall \text{x}}=  (x \in B)]\}$
\LAT
Propositiones 1...41 quoque subsistunt, si $a$, $b$... classes indicant; praeterea est:
\ENG
Propositions 1-41 also hold if $a,b,...$ indicate classes. In addition, we have:
\LAT
52. \hspace{0.67cm} $a$ \scalebox{0.7}[0.9]{$\in$} b . $\C$ . $b$ \scalebox{0.7}[0.9]{$\in$} \text{K}
\ENG
52. \hspace{0.67cm} $(a \in B) \rightarrow$ B is a \scriptsize{CLASS} \normalsize
\LAT
53. \hspace{0.67cm} $a$ \scalebox{0.7}[0.9]{$\in$} b . $\C$ . $b$ $-$ = $\abs$
\ENG
53. \hspace{0.67cm} $(a \in B) \rightarrow (B \not= \varnothing)$
\LAT
54. \hspace{0.67cm} $a$ \scalebox{0.7}[0.9]{$\in$} $b$ . $b$ = $c$ : $\C$ . $a$ \scalebox{0.7}[0.9]{$\in$} $c$
\ENG
54. \hspace{0.67cm} $[(a \in B) \wedge (B = C)] \rightarrow (a \in C)$
\LAT
55. \hspace{0.67cm} $a$ \scalebox{0.7}[0.9]{$\in$} $b$ . $b$ $\C$ $c$ : $\C$ . $a$ \scalebox{0.7}[0.9]{$\in$} $c$
\ENG
55. \hspace{0.67cm} $[(a \in B) \wedge (B \subset C)] \rightarrow (a \in C)$
\LAT
\quad Sit $s$ classis, et $k$ classis quae in $s$ contineatur; tunc dicimus $k$ esse individuum classis $s$, si $k$ ex uno tantum constat individuo. Itaque:
\ENG
\quad Let $\text{C}$ be a class, and $\text{C}_{subset}$ be a class which is contained in $\text{C}$; then we say that $\text{C}_{subset}$ is an individual of the class $\text{C}$, if $\text{C}_{subset}$ consists of only one individual. That is:
\LAT
56. \hspace{0.67cm} $s$ \scalebox{0.7}[0.9]{$\in$} K . $k$ $\C$ $s$ : $\C$ :: $k$ \scalebox{0.7}[0.9]{$\in$} $s$ . = $\pppNoSpace$ $k$ $-$ $=$ $\abs$ : $x$, $y$ \scalebox{0.7}[0.9]{$\in$} $k$ . $\C$\scalebox{0.7}{$x, y$}\thinspace . $x$ = $y$
\ENG
56. \hspace{0.67cm} $[\text{C}$ is a \scriptsize{CLASS} \normalsize $\wedge$ $(\text{C}_{subset} \subset \text{C})] \rightarrow \Big\{(\text{C}_{subset} \in \text{C}) = \big\{(\text{C}_{subset} \not= \varnothing) \wedge \{[(x,y) \in \text{C}_{subset}] \xrightarrow[\forall x,y]{} (x=y)\}\big\}\Big\}$
\end{translateTwoCol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{section}{V. The inverse}
\peanoHeadingSmall{V. De inversione.}
\ENG
\peanoHeadingSmall{V. The inverse.}

\raggedright
\fn{COMMENTARY: By inversion, Peano means going backwards from a proposition to a class.  It is not a mathematical inverse; it is ``set-builder notation''. }
\end{translateTwoCol}

\begin{translateTwoCol}
\quad Inversionis signum est $\scalebox{1.5}[0.8]{[\scalebox{0.2}{ \ }]}$, eiusque usum in sequenti numero explicabimus. Hic tantum casus particulares exponimus.
\ENG
\quad The sign of the inverse is $\{x | \ldots\}$, and we shall explain its use in the following section. Here we give some particular examples.
\LAT
\quad 1. Sit $a$ propositio, indeterminatum continens $x$; tunc scriptura $[x]$ \scalebox{0.7}[0.9]{$\in$} $a$, quae legitur \emph{ea x quibus a}, sive {\emph{solutiones}}, vel {\emph{radices}} conditionis $a$, classem significat individuis constitutam, quae conditioni $a$ satisfaciunt. Itaque:
\ENG
\quad 1. Let $a$ be a proposition containing the indeterminate $x$; then the expression $\{ x | a \}$, which is read {\emph{those x such that a}}, or {\emph{solutions}}, or {\emph{roots}} of the condition $a$, indicates the class consisting of individuals which satisfy the condition $a$. That is:
\LAT
57. \hspace{0.67cm} $a$ \scalebox{0.7}[0.9]{$\in$} \text{P} . $\C$ : $[x\scalebox{0.7}[0.9]{$\in$}]$ $a$ . \scalebox{0.7}[0.9]{$\in$} K
\ENG
57. \hspace{0.67cm} $a$ is a \scriptsize{proposition} \normalsize $\rightarrow \{x | a \}$ is a \scriptsize{CLASS} \normalsize 
\LAT
58. \hspace{0.67cm} $a$ \scalebox{0.7}[0.9]{$\in$} \text{K} . $\C$ $\pppNoSpace$ $[x\scalebox{0.7}[0.9]{$\in$}]$ . $x$ \scalebox{0.7}[0.9]{$\in$} $a$ : = $a$
\ENG
58. \hspace{0.67cm} $A$ is a \scriptsize{CLASS} \normalsize $\rightarrow \{ x | x \in A \} = A $
\LAT
59. \hspace{0.67cm} $a$ \scalebox{0.7}[0.9]{$\in$} \text{P} . $\C$ $\pppNoSpace$ $x$ \scalebox{0.7}[0.9]{$\in$} . $[x\scalebox{0.7}[0.9]{$\in$}]$ $a$ : = $a$
\ENG
59. \hspace{0.67cm}  $a$ is a \scriptsize{proposition} \normalsize $\rightarrow (x \in \{x | a \}) = a$
\LAT
\quad Sint $\alpha$, $\beta$ propositiones indeterminatum continentes $x$; erit:
\ENG
\quad Let $\alpha, \beta$ be propositions containing the indeterminate $x$. We will have:
\LAT
60. \hspace{0.67cm} $[x\scalebox{0.7}[0.9]{$\in$}]$ ($\alpha$ $\beta$) = ($[x\scalebox{0.7}[0.9]{$\in$}]$ $\alpha)$ $([x\scalebox{0.7}[0.9]{$\in$}]$ $\beta$)
\ENG
60. \hspace{0.67cm} $\{x | \alpha \wedge \beta\}=\{x|\alpha\} \cap \{x | \beta\}$
\LAT
61. \hspace{0.67cm} $[x\scalebox{0.7}[0.9]{$\in$}]$ $-$ $\alpha$ = $-$ $[x\scalebox{0.7}[0.9]{$\in$}]$ $\alpha$
\ENG
61. \hspace{0.67cm} $\{x | \neg \alpha \} = \overline{\{x | \alpha\}}$
\LAT
62. \hspace{0.67cm} $[x\scalebox{0.7}[0.9]{$\in$}]$ ($\alpha$ $\cup$ $\beta$) = $[x\scalebox{0.7}[0.9]{$\in$}]$ $\alpha$ $\cup$ $[x\scalebox{0.7}[0.9]{$\in$}]$ $\beta$
\ENG
62. \hspace{0.67cm} $\{ x | \alpha \vee \beta \} = \{ x | \alpha \} \cup \{ x | \beta \}$
\LAT
63. \hspace{0.67cm} $\alpha$ $\C$\scalebox{0.7}{$x$} $\beta$ . = . $[x\scalebox{0.7}[0.9]{$\in$}]$ $\alpha$  $\C$ $[x\scalebox{0.7}[0.9]{$\in$}]$ $\beta$
\ENG
63. \hspace{0.67cm} $\alpha \xrightarrow[\forall x]{} \beta = \{x | \alpha\} \subset \{x | \beta \}$
\LAT
64. \hspace{0.67cm} $\alpha$ =\scalebox{0.7}{$x$} $\beta$ . = . $[x\scalebox{0.7}[0.9]{$\in$}]$ $\alpha$ = $[x\scalebox{0.7}[0.9]{$\in$}]$ $\beta$
\ENG
64. \hspace{0.67cm} $(\alpha \underset{\forall \text{x}}= \beta) = (\{ x |  \alpha\} = \{x | \beta \} )$
\LAT
\quad 2. Sint $x,y$ entia quacumque; system ex ente $x$ et ex ente $y$ compositum ut novum ens consideramus, et signo $(x,y)$ indicamus; similiterque si entium numerus maior fit. Sit $\alpha$ propositio indeterminata continens $x,y$; tunc $[(x,y)\scalebox{0.7}[0.9]{$\in$}]$ $\alpha$ significat classem entibus $(x,y)$ constitutam, quae conditioni $\alpha$ satisfaciunt. Erit:
\ENG
\quad 2. Let $x,y$ be any entities. We shall consider the system composed of the entity $x$ and the entity $y$ as a new entity, and indicate it by the sign $(x,y)$; and similarly if the number of entities becomes larger. Let $\alpha$ be a proposition containing the indeterminate $x,y$; then $\{(x,y) | \alpha \}$ indicates the class of entities $(x,y)$ which satisfy the condition $\alpha$. We have:
\LAT
65. \hspace{0.67cm} $\alpha$ $\C$\scalebox{0.7}{$x, y$} $\beta$ . = . $[(x, y)\scalebox{0.7}[0.9]{$\in$}]$ $\alpha$  $\C$ $[(x, y)\scalebox{0.7}[0.9]{$\in$}]$ $\beta$
\ENG
65. \hspace{0.67cm} $\alpha \xrightarrow[\forall x,y]{} \beta = \{(x,y) | \alpha\} \subset \{(x,y) | \beta \}$
\LAT
66. \hspace{0.67cm} $[(x, y)\scalebox{0.7}[0.9]{$\in$}]$ $\alpha$ $-$ = $\abs$ . = $\pppNoSpace$ $[x\scalebox{0.7}[0.9]{$\in$}]$ . $[y\scalebox{0.7}[0.9]{$\in$}]$ $\alpha$ $-$ = $\abs$ : $-$ = $\abs$
\ENG
66. \hspace{0.67cm} $( \{ (x,y) | \alpha \} \neq \varnothing ) = ( \{ x | \{ y | \alpha \} \neq \varnothing \} \} \neq \varnothing )$
\LAT
\quad 3. Sit $x$ $\alpha$ $y$ relatio inter indeterminata $x$ et $y$ (ex. g. in logica relationes $x = y$, $x$ $- = y$, $x$ $\C$ $y$; in arithmetica $x < y$, $x > y$, etc). Tunc signo $[\scalebox{0.7}[0.9]{$\in$}$ $\alpha]$ $y$ ea $x$ indicamus, quae relationi $x$ $\alpha$ $y$ satisfaciunt. Commoditatis causa, loco $[\scalebox{0.7}[0.9]{$\in$}]$, signo $\reflectbox{\scalebox{0.7}[0.9]{$\in$}}$ utimur. Ita $\reflectbox{\scalebox{0.7}[0.9]{$\in$}}$ $\alpha$ $y$ . = : $[x$ $\scalebox{0.7}[0.9]{$\in$} ]$ . $x$ $\alpha$ $y$, et signum $\reflectbox{\scalebox{0.7}[0.9]{$\in$}}$ legitur \emph{qui}, vel \emph{quae}. Ex. gr. sit $y$ numerus; tunc $\reflectbox{\scalebox{0.7}[0.9]{$\in$}}$ $< y$ classem indicat numeris $x$ compositam qui conditioni $x<y$ satisfaciunt, scilicet, \emph{qui sunt minores} $y$, vel simpliciter \emph{minores} $y$. Similiter, quum signum D significet \emph{dividit}, vel \emph{est divisor}, formula $\reflectbox{\scalebox{0.7}[0.9]{$\in$}}$D significat \emph{qui dividunt} vel {divisores}. Deducitur $x$ $\scalebox{0.7}[0.9]{$\in$}$ $\reflectbox{\scalebox{0.7}[0.9]{$\in$}}$ $\alpha$ $y$ $=$ $x$ $\alpha$ $y$.
\ENG
\quad 3. Let $x$ $\alpha$ $y$ be a relation between the indeterminates $x$ and $y$ (eg. in logic, the relations $x = y$, $x \not= y$, $x \rightarrow y$; in arithmetic, $x < y$, $x > y$, ...). Then the sign $[\scalebox{0.7}[0.9]{$\in$}$ $\alpha]$ $y$ denotes the $x$ that satisfy the relation $x$ $\alpha$ $y$. For the sake of convenience, we use the sign $\reflectbox{\scalebox{0.7}[0.9]{$\in$}}$ instead of the sign $[\scalebox{0.7}[0.9]{$\in$}]$. Thus, $\reflectbox{\scalebox{0.7}[0.9]{$\in$}}$ $\alpha$ $y$  $= \{ x | x \alpha y \}$, and the sign $\reflectbox{\scalebox{0.7}[0.9]{$\in$}}$ is read \emph{the objects that}. For example, let $y$ be a number; then $\reflectbox{\scalebox{0.7}[0.9]{$\in$}}$ $< y$ denotes the class formed by the numbers $x$ that satisfy the condition $x<y$, that is, \emph{the objects that are smaller than} $y$, or simply \emph{the objects smaller than} $y$. Similarly, if the sign \emph{D} means \emph{divides} or \emph{is a divisor of}, the formula $\reflectbox{\scalebox{0.7}[0.9]{$\in$}}$D means \emph{the objects that divide} or \emph{the divisors}. It follows that $x$ $\scalebox{0.7}[0.9]{$\in$}$ $(\reflectbox{\scalebox{0.7}[0.9]{$\in$}}$ $\alpha$ $y)$ $=$ $x$ $\alpha$ $y$.
\LAT
\quad 4. Sit $\alpha$ formula indeterminate continens $x$. Tunc scriptura $x' [x] \alpha$, quae legitur $x'$ \emph{loco} $x$ \emph{in} $\alpha$ \emph{substituto}, formulam indicat quae obtinetur si in $\alpha$, loco $x$, $x'$ legimus. Deducitur $x [x] \alpha = \alpha$.
\ENG
\quad 4. Let $\alpha$ be a formula containing the indeterminate $x$. Then the expression $\alpha [ x := x' ]$, which is read $x'$ \emph{being substituted for} $x$ \emph{in} $\alpha$, denotes the formula obtained if, in $\alpha$, we read $x'$ for $x$. It follows that $\alpha[x := x] = \alpha$.
\LAT
\quad 5. Sit $\alpha$ formula, quae indeterminata $x,y,...$ continet. Tunc
\ENG
\quad 5. Let $\alpha$ be a formula that contains the indeterminates $x,y,...$ Then
\LAT
\hspace{1.06cm} $(x',y',...)$ $[x,y,...]$ $\alpha$,
\ENG
\hspace{1.06cm} $\alpha[x:=x', y:=y', \ldots]$,
\end{translateTwoCol}

\peanoPage{XIII} % page-number XIII

\begin{translateTwoCol}
quae legitur $x'y',...$ \emph{loco} $x,y,...$ \emph{in} $\alpha$ \emph{substitutis}, formulam indicat quae obtinetur si in $\alpha$ loco $x,y,...$, litterae $x'y',...$ scribantur. Deducitur $(x,y)$ $[x,y]$ $\alpha = \alpha$.
\ENG
which is read $x',y',...$ \emph{being substituted for} $x,y,...$ in $\alpha$, denotes the formula obtained if, in $\alpha$, the letters $x',y',...$ are written for $x,y,...$ It follows that $\alpha[x:=x,y:=y] = \alpha$.
\end{translateTwoCol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{section}{VI. Functions}
\peanoHeadingSmall{VI. De functionibus.}
\ENG
\peanoHeadingSmall{VI. Functions.}

\raggedright
\fn{COMMENTARY: Peano has a very structural or syntactical version of a function.  There is no parameter.  The function is text like ``2 + `` or ``+ 3'' where if you append a number to the start or end, you have a valid expression.}
\end{translateTwoCol}

\begin{translateTwoCol}
Logicae notationes quae praecedunt exprimendae cuilibet arithmeticae propositioni sufficiunt, iisdemque tantum utimur. Hic notationes alias nonnullas breviter explicamus, quae utiles fieri possunt.
\ENG
The symbols of logic introduced above suffice to express any proposition of arithmetic, and we shall only these.We explain here briefly some other symbols that may be useful.
\LAT
Sit $s$ quaedam classis; supponimus aequalitatem inter entia systematis $s$ definitam, quae conditionibus satisfaciat:
\ENG
Let $S$ be a class; we assume that equality is defined between the objects of the system $s$ so as to satisfy the conditions:
\LAT
\hspace{1.06cm} $a=a$
\ENG
\hspace{1.06cm} $a=a$.
\LAT
\hspace{1.06cm} $a=b$ $.$ $=$ $.$ $b=a$
\ENG
\hspace{1.06cm} $(a=b)$ $=$ $(b=a)$
\LAT
\hspace{1.06cm} $a=b$ . $b=c$ $:$ $\C$ $.$ $a=c$
\ENG
\hspace{1.06cm} $[(a=b)$ $\wedge$ $(b=c)]$ $\rightarrow$ $a=c$
\LAT
Sit $\upvarphi$ signum, sive signorum aggregatus, ita ut si $x$ est ens classis $s$, scriptura $\upvarphi$ $x$ novum indicet ens; supponimus quoque aequalitatem inter entia $\upvarphi$ $x$ definitam; et si $x$ et $y$ sunt entia classis $s$, et est $x=y$, supponimus deduci posse $\upvarphi$ $x$ $=$ $\upvarphi$ $y$. Tunc signum $\upvarphi$ dicitur esse \emph{functionis praesignum in classi} $s$, et scribemus $\upvarphi$ $\reflectbox{\scalebox{0.7}[0.9]{$\in$}}$ $F'$ $s$.
\ENG
Let $\upvarphi$ be a sign or an aggregate of signs such that, if $x$ is an object of the class $S$, the expression $\upvarphi$ $x$ denotes a new object; we assume also that equality is defined between the objects $\upvarphi$ $x$; further, if $x$ and $y$ are objects of the class $S$ and if $x = y$, we assume it is possible to deduce $\upvarphi$ $x = \upvarphi$ $y$. Then the sign $\upvarphi$ is said to be a \emph{function presign in the class S}, and we write $\upvarphi$ $\reflectbox{\scalebox{0.7}[0.9]{$\in$}}$ $F'$ $s$.
\LAT
\hspace{1.06cm} $s$ $\scalebox{0.7}[0.9]{$\in$}$ $K$ $.$ $\C$ $:\hspace{0.1em}:$ $\upvarphi$ $F'$ $s$ $.$ $=$ $\pppNoSpace$ $x,y$ $\scalebox{0.7}[0.9]{$\in$}$ $s$ $.$ $x$ $=$ $y$ $:$ $\C_{x,y}$ $.$ $\upvarphi x$ $=$ $\upvarphi y$
\ENG
...
\LAT
Verum si, cum sit $x$ quodlibet ens classis $s$, scriptura $x\upvarphi$ novum indicet ens, et, ex, $x$ $=$ $y$ deducitur $x\upvarphi$ $=$ $y\upvarphi$, tunc dicimus $\upvarphi$ esse \emph{functionis postsignum in classi} $s$ et scribemus $\upvarphi$ $\scalebox{0.7}[0.9]{$\in$}$ $s'$F.
\ENG
If, $x$ being any object of the class $S$, the expression $x\upvarphi$ denotes a new object and $x\upvarphi = y\upvarphi$ follows from $x=y$, then we say that $\upvarphi$ is a \emph{function postsign in the class S}, and we write $\upvarphi$ $\scalebox{0.7}[0.9]{$\in$}$ $s'$F.
\LAT
\hspace{1.06cm} $s$ $\scalebox{0.7}[0.9]{$\in$}$ $K$ $.$ $\C$ $:\hspace{0.1em}:$ $\upvarphi$ $s$ $'F$ $.$ $=$ $\pppNoSpace$ $x,y$ $\scalebox{0.7}[0.9]{$\in$}$ $s$ $.$ $x$ $=$ $y$ $:$ $\C_{x,y}$ $.$ $x\upvarphi$ $=$ $y\upvarphi$
\ENG
...
\LAT
\emph{Exempla}. Sit $a$ numerus; tunc $a$ $+$ est functionis praesignum in numerorum classe, et $+$ $a$ est functionis postsignum; quicumque enim est numerus $x$, formulae $a$ $+$ $x$ et $x$ $+$ $a$ novos indicant numeros, et ex $x$ $=$ $y$ deducitur $a$ $+$ $x$ $=$ $a$ $+$ $y$, et $x$ $+$ $a$ $=$ $y$ $+$ $a$. Itaque
\ENG
\emph{Examples}. Let $a$ be a number; then $a+$ is a function presign in the class of numbers, and $+a$ is a function postsign; for any number $x$, formulas $a+x$ and $x+a$ denote new numbers; $a+x=a+y$ and $x+a=y+a$ follow from $x=y$. Thus
\LAT
\hspace{1.06cm} $a$ $\scalebox{0.7}[0.9]{$\in$} \N . \C : a + . \in .$ $F' \N$
\ENG
\hspace{1.06cm}
\LAT
\hspace{1.06cm} $a$ $\scalebox{0.7}[0.9]{$\in$} \N .$ $\C$ $:$ $+$ $a$ $.$ $\scalebox{0.7}[0.9]{$\in$}$ $.$ $N$ $'F$
\ENG
...
\LAT
\hfill
\ENG
\hfill
\LAT
Sit $\upvarphi$ functionis praesignum in classe $s$. Tunc $[\upvarphi] y$ classem significat iis $x$ constitutam, quae conditioni $\upvarphi x$ $=$ $y$ satisfaciunt; scilicet:
\ENG
Let $\upvarphi$ be a function presign in the class $S$. Then $[\upvarphi] y$ denotes the class composed of $x$ that satisfy the condition $\upvarphi x = y$; that is,
\LAT
\emph{Def}. \hspace{0.25cm} $s$ $\scalebox{0.7}[0.9]{$\in$}$ K $.$ $\upvarphi$ $\scalebox{0.7}[0.9]{$\in$}$ $F'$ $s$ $:$ $\C$ $:$ $[\upvarphi]y$ $.$ $=$ $.$ $[x \scalebox{0.7}[0.9]{$\in$}]$ $(\upvarphi x$ $=$ $y)$
\ENG
\emph{Def}. \hspace{0.25cm}
\end{translateTwoCol}

\peanoPage{XIV} % page-number XIV

\begin{translateTwoCol}
Classis $[\upvarphi] y$ vel unum vel plura, vel etiam nullum individuum continere potest. Erit:
\ENG
The class $[\upvarphi] y$ may contain one or several individuals, or none at all. We have:
\LAT
\hspace{1.06cm} $s$ $\scalebox{0.7}[0.9]{$\in$}$ K $.$ $\upvarphi$ $\scalebox{0.7}[0.9]{$\in$}$ $F'$ $s$ $:$ $\C$ $:$ $y$ $=$ $\upvarphi$ $x$ $.$ $=$ $.$ $x$ $\scalebox{0.7}[0.9]{$\in$}$ $[\upvarphi] y$
\ENG
...
\LAT
Si vero $\upvarphi y$ uno tantum constat individuo, erit $y$ $=$ $\upvarphi x$ $.$ $=$ $.$ $x$ $=$ $[\upvarphi] y$
\ENG
But if $\upvarphi y$ consists of just one individual, we have $y$ $=$ $\upvarphi x$ $.$ $=$ $.$ $x$ $=$ $[\upvarphi] y$
\LAT
Sit $\upvarphi$ functions postsignum; similiter ponimus:
\ENG
Let $\upvarphi$ be a function postsign; we write similarly:
\LAT
\hspace{1.06cm} $s$ $\scalebox{0.7}[0.9]{$\in$}$ K $.$ $\upvarphi$ $\scalebox{0.7}[0.9]{$\in$}$ $s$ $'F$ $:$ $\C$ $\pppNoSpace$ $y$ $| \upvarphi |$ $=$ $| x\scalebox{0.7}[0.9]{$\in$} |$ $(x \upvarphi = y)$.
\ENG
...
\LAT
Signum $[$ $]$ dicitur \emph{inversionis signum}, eiusque usus nonullos in logica iam exposuimus. Nam si $\alpha$ est propositio indeterminatum continens $x$, atque $A$ est classis individuis $x$ composita quae conditioni $\alpha$ satisfaciunt, erit $x$ $\scalebox{0.7}[0.9]{$\in$}$ $a$ $.$ $=$ $\alpha$, tunc $a$ $=$ $[x$\scalebox{0.7}[0.9]{$\in$}$]$ $\alpha$, ut in V, i.
\ENG
The sign $\{ x | \ldots \}$ is called \emph{inversion sign}, and we have already presented some of its uses in logic. If $\alpha$ is a proposition containing the indeterminate $x$ and $A$ is a class composed of the individuals $x$ that satisfy the condition $\alpha$, we have $x$ $\scalebox{0.7}[0.9]{$\in$}$ $A$ $=$ $\alpha$, and then $A$ $=$ $\{ x | \alpha \}$, as in V, i.
\LAT
Sit $\alpha$ formula indeterminate continens $x$, sitque $\upvarphi$ functionis praesignum, quod litterae $x$ praepositum, formulam $\alpha$ gignat; scilicet sit $\alpha$ $=$ $\upvarphi$ $x$; tunc erit $\upvarphi$ $=$ $\alpha [x]$, et si $x'$ est novum ens, erit $\upvarphi x'$ $=$ $\alpha [x] x'$, scilicet, si $\alpha$ est formula indeterminatum continens $x$, tunc $\alpha [x] x'$ significat id quod obtinetur si in $\alpha$, loco $x, x'$ ponatur.
\ENG
Let $\alpha$ be a formula containing the indeterminate $x$ and let $\upvarphi$ be a function presign that yields the formula $\alpha$ when written before the letter $x$; that is, let $\alpha = \upvarphi x$; then we have $\upvarphi$ $=$ $\alpha [x := ? ]$, and if $x'$ is a new object, we have $\upvarphi x'$ $=$ $\alpha [x := x']$, that is, if $\alpha$ is a formula containing the indeterminate $x$, then $\alpha[x := x']$ means what is obtained when, in $\alpha$, we put $x'$ for $x$.
% TODO: ADD COMMENT THAT THE ABOVE DOESN'T WORK WITH PEANO'S SYNTACTIC TRICK
\LAT
Similiter, sit $\alpha$ formula indeterminate continens $x$, sitque $\upvarphi$ functionis postsignum, ut $x \upvarphi = \alpha$; deducitur $\upvarphi = [x] \alpha$; tunc, si $x'$ est novum ens, erit $x' \upvarphi = x' [x] \alpha$, scilicet $x' [x] \alpha$ rursum indicat id quod obtinetur si in $\alpha$, loco $x, x'$ legatur, ut in V, 4.
\ENG
Similarly, let $\alpha$ be a formula containing the indeterminate $x$ and let $\upvarphi$ be a function postsign, such that $x\upvarphi = \alpha$; it follows that $\upvarphi = \alpha[x := ?]$. Then, if $x'$ is a new object, have $x' \upvarphi = \alpha[x := x']$; that is, $x' [x] \alpha$ again denotes what is obtained, when, in $\alpha$, read $x'$ for $x$, as in V, 4..
% TODO: ADD COMMENT SAYING THIS DOESN'T WORK.  HOW DO BOTH WORK IN PEANO'S NOTATION???!!!
\LAT
Alios quoque usus in logica signum $[ ]$ habere potest, quos breviter esponimus, quum ipsis non utamur. Sint $a$ et $b$ duae classes; tunc $[a \cap ]b$ sive $b[ \cap a]$ classes indicat $x$, quae conditioni $b = a \cap x$, sive $b = x \cap a$ satisfaciunt. Si $b$ in $a$ non continetur, nulla classis huic conditioni satifacit; si $b$ in $a$ continetur, signum $b [ \cap a ]$ omnes indicat classes quae $b$ continent atque in $b \cup - a$ continentur.
\ENG
The sign $[ ]$ can have other uses in logic, which we present only briefly, since we shall not use it in these ways. Let $A$ and $B$ be two classes; then $[A \cap ]B$ or $B[ \cap A]$ denotes the classes $X$ that satisfy the condition $B = A \cap X$, or $B = X \cap A$. If $B$ is not contained in $A$, no class satisfies this condition; if $B$ is contained in $A$, the sign $B [ \cap A ]$ denotes all classes that contain $B$ and are contained in $B \cup  \overline{A}$.
\LAT
In Arithmetica, sint $a, b$ numeri; tunc $[b + a]$ sive $[a +] b$ numerum indicat $x$, qui conditioni $b = x + a$, sive $b = a + x$ satisfacit, nempe $b - a$. Similiter erit $b [ \times a ] = [a \times ] b = b / a$. Et in analysi hoc signum usuvenire potest; itaque
\ENG
In arithmetic, let $a$ and $b$ be numbers; then $[b + a]$ or $[a +] b$ denotes the number $x$ that satisfies the condition $b = x + a$, or $b = a + x$, that is $b - a$. Similarly we have $b [ \times a ] = [a \times ] b = b / a$. This sign can even find a use in analysis; thus
\end{translateTwoCol}

\columnratio{0.25, 0.25, 0.25, 0.25}
\begin{paracol}{4}
\hspace{0.54cm} $y = sin$ $x$ $.$ = $.$ $x \scalebox{0.7}[0.9]{$\in$}$ $[sin]$ $y$
\s
(loco $x = arc$ $sin$ $y).$
\s
...
\s
...
\s
\hspace{0.54cm} $d F (x) = f(x) dx$ $.$ $=$ $.$ $F$ $(x)\scalebox{0.7}[0.9]{$\in$}$ $[d]$ $f(x)dx$
\s
(loco $F(x) = \int{f(x)dx}).$ 
\s
...
\s
...
\end{paracol}

\begin{translateTwoCol}
\hfill
\ENG
\hfill
\LAT
Sit rursum $\upvarphi$ functionis praesignum in classi $s$, sitque $k$ classis
\ENG
Let $\upvarphi$ again be a function presign in a class $S$ and let $C$ be a class
\end{translateTwoCol}

\peanoPage{XV} % page-number XV

\begin{translateTwoCol}
in $s$ contenta; tunc $\upvarphi k$ classem indicat omnibus $\upvarphi x$ compositam, ubi $x$ sunt entia classis $k$; scilicet
\ENG
contained in $S$; then $\upvarphi C$ denotes the class consisting of all $\upvarphi x$, where the $x$ are the objects of class $C$; that is
\LAT
\emph{Def.} \hspace{0.25cm} $s$ $\scalebox{0.7}[0.9]{$\in$}$ K $.$ $k$ $\scalebox{0.7}[0.9]{$\in$}$ K $.$ $k$ $\C$ $s$ $.$ $\upvarphi$ $\scalebox{0.7}[0.9]{$\in$}$ $F'$ $s$ $:$ $\C$ $.$ $\upvarphi$ $k$ $=$ $[y $\scalebox{0.7}[0.9]{$\in$}$]$ $(k$ $.$ $[\upvarphi]y$ $:$ $-$ $=$ $\abs$)
\ENG
\emph{Def}. \hspace{0.25cm}  %TODO
\LAT
Sive \hspace{0.25cm} $s$ $\scalebox{0.7}[0.9]{$\in$}$ K $.$ $k$ $\scalebox{0.7}[0.9]{$\in$}$ K $.$ $k$ $\C$ $s$ $.$ $\upvarphi$ $\scalebox{0.7}[0.9]{$\in$}$ $F'$ $s$ $:$ $\C$ $.$ $\upvarphi$ $k$ $=$ $[y $\scalebox{0.7}[0.9]{$\in$}$]$ $([x\scalebox{0.7}[0.9]{$\in$}]$ $:$ $x$ $\scalebox{0.7}[0.9]{$\in$}$ $k$ $.$ $[\upvarphi]x$ $=$ $y$ $\pppNoSpace$ $-$ $=$ $\abs$)
\ENG
Or \hspace{0.25cm}  %TODO
\LAT
\emph{Def.} \hspace{0.25cm} $s$ $\scalebox{0.7}[0.9]{$\in$}$ K $.$ $k$ $\scalebox{0.7}[0.9]{$\in$}$ K $.$ $k$ $\C$ $s$ $.$ $\upvarphi$ $\scalebox{0.7}[0.9]{$\in$}$ $s$ $'F$ $:$ $\C$ $.$ $k$ $\upvarphi$ $=$ $[y $\scalebox{0.7}[0.9]{$\in$}$]$ $(k$ $.$ $y$ $[\upvarphi]$ $:$ $-$ $=$ $\abs$)
\ENG
\emph{Def.} \hspace{0.25cm} %TODO
\LAT
Itaque, si $\upvarphi$ $\scalebox{0.7}[0.9]{$\in$}$ $F' s$, tunc $\upvarphi$ $s$ classem indicat omnibus $\upvarphi$ $x$ constitutam, ubi $x$ sint entia classis $s$. Erit:
\ENG
Thus, if $\upvarphi$ $\scalebox{0.7}[0.9]{$\in$}$ $F' S$, then $\upvarphi$ $s$ denotes the class composed of all $\upvarphi$ $x$, where the $x$ are objects of the class $S$. We have:
\LAT
\hspace{1.06cm} $s \in K$ . $\upvarphi \in F' s$ . $y \in \upvarphi s : \C : \upvarphi [\upvarphi] y = y$
\ENG
... %TODO
\LAT
\hspace{1.06cm} $s \in K$ . $a, b \in K$ . $a$ $\C$ $s$ . $b$ $\C$ $s$ . $\upvarphi \in F' s : \C$ . $\upvarphi (a \cup b) = (\upvarphi a) \cup (\upvarphi b)$
\ENG
... %TODO
\LAT
\hspace{1.06cm} $s \in K$ . $\upvarphi \in F' s : \C$ . $\upvarphi \abs = \abs$
\ENG
... %TODO
\LAT
\hspace{1.06cm} $s \in K$ . $a,b \in K$ . $b$ $\C$ $s$ . $a$ $\C$ $b$ . $\upvarphi \in F' s : \C$ . $\upvarphi a$ $\C$ $\upvarphi b$
\ENG
... %TODO
\LAT
\hspace{1.06cm} $s \in K$ . $a,b \in K$ . $a$ $\C$ $s$ . $b$ $\C$ $s$ . $\upvarphi \in F' s : \C$ . $\upvarphi (ab)$ $\C$ $(\upvarphi a)(\upvarphi b)$
\ENG
... %TODO
\LAT
Sit $a$ quaedam classis; tunc $a \cap K$, sive $K \cap a$, sive $K a$, classes omnes indicat formae $a \cap x$, sive $x \cap a$, $x a$, ubi $x$ est classis quacumque; scilicet $K a$ indicat classes quae in $a$ continentur. Formula $x \in K a$ idem significat quod $x \in K$ . $x$ $\C$ $a$. Hac conventione quandoque utimur; ita $K N$ isgnificat \emph{numerorum classem}.
\ENG
Let $a$ be a class; then $a \cap K$, or $K \cap a$, or $K a$, denotes all classes of the form $a \cap x$, or $x \cap a$, $x a$, where $x$ is any class; that is $K a$ denotes the classes that are contained in $a$. The formula $x \in K a$ means the same as  $x \in K$ . $x$ $\C$ $a$. We shall sometimes use this convention; thus $K N$ means \emph{a class of numbers}.
\LAT
Similiter, si $a$ est classis, $K \cup a$ indicat classes quae $a$ continent. Sit $a$ numerus; tunc $a + N$, sive $N + a$, \emph{numeros} indicat \emph{numero a maiores}; $a \times N$, sive $N \times a$, sive $N a$ indicat \emph{multiplices numeri a}; $a^N$ indicat \emph{potestas numeri} $a$; $N^2$, $N^3$, ... indicat \emph{numeros quadratos}, vel \emph{numeros cubos},...
\ENG
Similarly, if $A$ is a class, $K \cup A$ indicates the classes that contain $A$. Let $a$ be a number; then $a + N$, or $N + a$, denotes \emph{the numbers greater than the number} $a$; $a \times N$, or $N \times a$, or $N a$ denotes \emph{the multiples of the number} $a$; $a^N$ denotes \emph{the powers of the number} $a$; $N^2$, $N^3$, ... denote \emph{the squares}, \emph{the cubes},...
\LAT
\hfill
\ENG
\hfill
\LAT
Functional signorum aequalitatem, productum, potestas, ita definire licet:
\ENG
Equality, product and powers can be defined thus for function signs:
\LAT
\emph{Def.} \hspace{0.25cm} $s \in K$ . $\varphi, \psi \in F' s :$ $\C$ $\pppNoSpace$ $\varphi = \psi$ : $=$ : $x \in s$ . $\C$ . $\varphi x = \psi x$
\ENG
\emph{Def.} \hspace{0.25cm}  %TODO
\LAT
\emph{Def.} \hspace{0.25cm} $s \in K$ . $\varphi \in F' s$ $.$ $\psi \in F' \varphi s$ . $x \in s$ : $\C$ . $\psi \varphi x = \psi (\varphi x)$
\ENG
\emph{Def.} \hspace{0.25cm}  %TODO
\LAT
Itaque, in definitionis hypothesi, erit $\psi \varphi$ novum functionis praesignum; idque \emph{productum signorum} $\psi$ \emph{et} $\varphi$ vocatur.
\ENG
Thus, if we assume this definition, we have the new function presign $\psi \varphi$; it is called the \emph{product of the signs} $\psi$ and $\varphi$.
\LAT
Similiterque, si $\varphi$, $\psi$ sunt functionis postsigna.
\ENG
Similarly if $\varphi$, $\psi$ are function postsigns.
\LAT
Haec valet propositio:
\ENG
The following proposition holds:
\LAT
\hspace{1.06cm} $s \in K$ . $\varphi \in F' s$ . $\varphi s$ $\C$ $s : \C : \varphi \varphi s$ $\C$ $s$ . $\varphi \varphi \varphi s$ $\C$ $s$. etc.
\ENG
...
\LAT
Funcitones $\varphi \varphi, \varphi \varphi \varphi,...$ \emph{iteraiae} vocantur, et communiter signis $\varphi^2, \varphi^3,...$ indicantur, ut operationis $\varphi$ potestates.
\ENG
The functions $\varphi \varphi, \varphi \varphi \varphi,...$ are said to be \emph{iterated} and are generally denoted by the signs $\varphi^2, \varphi^3,...$ as powers of the operation $\varphi$.
\end{translateTwoCol}

\peanoPage{XVI} % page-number XVI

\begin{translateTwoCol}
Si vero $\varphi$ est functionis postsignum, ha faciliori notatione, absque ambiguitate, uti licet:
\ENG
But if $\varphi$ is a function postsign, we can use the following more convenient notation without ambiguity:
\LAT
\emph{Def.} \hspace{0.25cm} $s \in K$ . $\varphi \in s 'F$ . $s \varphi$ $\C$ $s : \C : \varphi 1 = \varphi$ . $\varphi 2 = \varphi \varphi$ . $\varphi 3 = \varphi \varphi \varphi$. etc.
\ENG
\emph{Def.} \hspace{0.25cm}  %TODO
\LAT
In definitionis hypothesi, si $m, n \in N$, erit $\varphi$ $(m+n)$ $=$ $(\varphi m)(\varphi n)$; $(\varphi m)n$ $=$ $\varphi (m n)$
\ENG
Assuming this definition, if $m, n \in N$, we have $\varphi$ $(m+n)$ $=$ $(\varphi m)(\varphi n)$; $(\varphi m)n$ $=$ $\varphi (m n)$
\LAT
Si hac definitione in Arithmetica utimur, haec invenimus. Numerum qui sequitur numerum $a$ signo faciliori $a +$ indicare possumus; tunc $a + 1, a+2,...$ et, si $b$ est numerus, $a + b$, sensum habent $a +, a + +,...$ quod a definitione in \S 1 patet. Propositionem 6 in \S 1 scribere possumus $N +$ $\C$ $N$. Si $a, b, c$ sunt numeri, tunc $a : + b$ . $c$ significat $a + b c$, et $a : \times b$ . $c$ significat $a b^c$.
\ENG
If we use this definition in arithmetice, we obtain the following. We can denote the number that follows the number $a$ by the more convenient sign $a+$; then $a + 1$, $a + 2$,..., and, if $b$ is a number, $a + b$, have the meaning of $a +$, $a ++$,..., which is clear from the definition in \S 1 below. Proposition 6 in \S 1 can be written $N +$ $\C$ $N$. If $a, b, c$ are numbers, then $a : + b$ . $c$ means $a + b c$, and $a : \times b$ . $c$ means $a b^c$.
\LAT
Multi aliis proprietatibus gaudent functionem signa, praesertim si conditioni satisfaciunt: $\varphi x = \varphi y$ . $\C$ . $x = y$. Functionis signum quod huic conditioni satisfacit vocatur a clarrissimo Dedekind \emph{simile} (hnliche Abbildung).
\ENG
Function signs possess many other properties, especially if they satisfy the condition: $\varphi x = \varphi y$ . $\C$ . $x = y$. A function sign that satisfies this condition is called \emph{equivalent}\endnote{The 2 other translation, mentioned at the beginning of this current document, translated "hnlich" literally to "similar", instead of "equivalent". However, additional information can be found in a footnote of the first translation: \\

\quad \emph{"Today "similar" has another meaning and instead we would say "equivalent"."} \\

G. Peano, (1889), ``The principles of arithmetic presented by a new method" in: J. van Heijenoort (ed.), \emph{From Frege to G\"odel. A source book in mathematical logic. 1879-1931}, Cambridge: Harvard University Press, 1967, p. 93.} by Dedekind (hnliche Abbildung).
\LAT
Sed his exponendis locus deest.
\ENG
But we lack the space to present these properties.
\end{translateTwoCol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{section}{Remarks}
\peanoHeadingSmall{Declarationes.}
\ENG
\peanoHeadingSmall{Remarks.}
\end{translateTwoCol}

\begin{translateTwoCol}
\emph{Defenitio}, vel breviter \emph{Def.} est propositio formam habens $x = a$, sive $\alpha$ $\C$ . $x = a$, ubi $\alpha$ est signorum aggregatus sensum habens notum; $x$ est signum, vel signorum aggregatus significatione adhuc carnes; $\alpha$ vero est conditio sub qua definitio datur.
\ENG
A \emph{Definition}, or \emph{Def.} for short, is a proposition of the form $x=a$ or $\alpha$ $\C$ . $x = a$, where $a$ is an aggregate of signs having a known meaning, $x$ is a sign or an aggregate of signs, hitherto without meaning, and $\alpha$ is the condition under which the definition is given.
\LAT
\emph{Theorema}, (Theor. vel Th) est propositio quae demonstratur. Si theorema formam habet $\alpha$ $\C$ $\beta$, ubi $\alpha$ et $\beta$ sunt propositiones, tunc $\alpha$ dicitur \emph{Hypothesis} (Hyp. vel breviter Hp.), $\beta$ vero \emph{Thesis} (Thes. vel Ts.). Hyp. ac Ts. a Theorematis forma pendent; nam si loco $\alpha$ $\C$ $\beta$ scribemus $- \beta$ $\C$ $- \alpha$, erit $- \beta$ Hp, et $- \alpha$ Ts.; si vero scribemus $\alpha - \beta = \abs$, Hp. ac Ts. absunt.
\ENG
A \emph{theorem} (Theor. or Th.) is aproposition that is \emph{proved}. If a theorem has the form $\alpha$ $\C$ $\beta$, where $\alpha$ and $\beta$ are propositions, then $\alpha$ is called the \emph{hypothesis} (Hyp., or even shorter, Hp.) and $\beta$ the \emph{thesis} (Thes. or Ts.). Hyp. and Ts. depend on the form of the theorem; in fact, if we write $\alpha$ $\C$ $\beta$ instead of $- \beta$ $\C$ $- \alpha$, then $- \beta$ is the Hp, abd $- \alpha$ the Ts.; if we write $\alpha - \beta = \abs$, Hp. and Ts. do not exist.
\LAT
In quolibet $\S$ signum P quod quidam numerus sequatur, propositionem indicat eiusdem $\S$ hoc numero signatam. Logicae propositiones indicantur signo L et propositiones numero.
\ENG
In any $\S$ below, the sign P followed by a number denotes the proposition indicated by that number in the same $\S$. Propositions of logic are indicated by the sign L and the number of the proposition.
\LAT
Formulae quae in una linea non continentur, in altera linea, nullo interposito signo, sequuntur.
\ENG
Formulas that do not fit on one line are continued on the next line without any intervening sign.
\end{translateTwoCol}

\pagebreak

\peanoPage{1} % page-number 1

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{chapter}{The principles of arithmetic}
\peanoHeadingLarge{Arithmetices Principia.}
\ENG
\peanoHeadingLarge{The Principles of Arithmetic.}
\LAT
\phantomsection
\addcontentsline{toc}{section}{\S1. Numbers and addition}
\peanoHeadingMedium{\S1. De numeris et de additione.}
\ENG
\peanoHeadingMedium{\S1. Numbers and addition.}
\LAT
\phantomsection
\addcontentsline{toc}{subsection}{Explanations}
\peanoHeadingSmall{Explicationes.}
\ENG
\peanoHeadingSmall{Explanations.}

\raggedright
\fn{COMMENTARY: Peano starts his natural numbers at 1; most modern versions start at 0.  Also, while Peano calls it ``successor'' (``sequens''), he does not use the modern convention of using ``S'' for it.}
\end{translateTwoCol}

\columnratio{0.1, 0.05, 0.1, 0.25, 0.1, 0.05, 0.1, 0.25}
\begin{paracol}{8}
\centering Signo \s $\N$ \s significatur \s \raggedright \emph{numerus (integer positivus)}.
\s
\centering The sign \s $\N$ \s means \s \raggedright \emph{number (positive integer)}.
\s*
\centering $\dittoMarkLatin$ \s $1$ \s $\dittoMarkLatin$ \s \raggedright \emph{unitas}.
\s
\centering \dittoMarkEnglish \s $1$ \s \dittoMarkEnglish \s \raggedright \emph{unity}.
\s*
\centering $\dittoMarkLatin$ \s $a + 1$ \s $\dittoMarkLatin$ \s \raggedright \emph{sequens} $a$, sive $a$ \emph{plus} $1$.
\s
\centering \dittoMarkEnglish \s $a + 1$ \s \dittoMarkEnglish \s \raggedright \emph{the successor of} $a$, or $a$ \emph{plus} $1$.
\s*
\centering $\dittoMarkLatin$ \s $=$  \s $\dittoMarkLatin$ \s \raggedright \emph{est aequalis}. Hoc ut novum signum considerandum est, etsi logicae signi figuram habeat.
\s
\centering \dittoMarkEnglish \s $=$ \s \dittoMarkEnglish \s \raggedright \emph{is equal to}. This must be considered as a new sign, although it has the appearance of a sign of logic.
\end{paracol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Axioms}
\peanoHeadingSmall{Axiomata.}
\ENG
\peanoHeadingSmall{Axioms.}
\end{translateTwoCol}

\begin{translateTwoCol}
1. \hspace{0.6cm} $1 \in N$
\ENG
...
\LAT
2. \hspace{0.6cm} $a \in \N \p \C \p a = a$
\ENG
...
\LAT
3. \hspace{0.6cm} $a,b,c \in \N \p \C : a = b \p = \p b = a$
\ENG
...
\LAT
4. \hspace{0.6cm} $a,b \in \N \p \C \ppp a = b \p b = c : \C \p a = c$
\ENG
...
\LAT
5. \hspace{0.6cm} $a = b \p b \in \N : \C \p a \in N$
\ENG
...
\LAT
6. \hspace{0.6cm} $a \in \N \p \C \p a + 1 \in N$
\ENG
...
\LAT
7. \hspace{0.6cm} $a, b \in \N \p \C : a = b \p = \p a + 1 = b + 1$
\ENG
...
\LAT
8. \hspace{0.6cm} $a \in \N \p \C \p a + 1 \no = 1$
\ENG
...
\LAT
9. \hspace{0.6cm} $k \in K \ppp 1 \in k \ppp x \in \N \p x \in k : \C_x \p x + 1 \in k : : \C \p \N \C k$
\ENG
...
\end{translateTwoCol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Definitions}
\peanoHeadingSmall{Definitiones.}
\ENG
\peanoHeadingSmall{Definitions.}
\end{translateTwoCol}

\begin{translateTwoCol}
10. \hspace{0.67cm} $2 = 1 + 1;$ $3 = 2 + 1;$ $4 = 3 +1;$ etc.
\ENG
...
\end{translateTwoCol}

\peanoPage{2} % page-number 2

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Theorems}
\peanoHeadingSmall{Theoremata.}
\ENG
\peanoHeadingSmall{Theorems.}
\end{translateTwoCol}

\begin{translateTwoCol}
11. \hspace{0.67cm} $2 \in N$.
\ENG
...
\LAT
\vspace{1em}
\emph{Demonstratio:}
\vspace{1em}
\ENG
\vspace{1em}
\emph{Proof:}
\vspace{1em}
\end{translateTwoCol}

\columnratio{0.26, 0.18, 0.06, 0.27, 0.18, 0.06}
\begin{paracol}{6}
\raggedright
\setlength{\columnsep}{1em}
$P 1 \p \C :$ \s $1 \in N$ \s (1)
\s
.... \s .... \s (1)
\s*
$1 [a] (P6) \p \C :$ \s $1 \in \N \p \C \p 1 + 1 \in N$ \s (2)
\s
... \s ... \s (2)
\s*
$(1)(2) \p \C :$ \s $1 + 1 \in N$ \s (3)
\s
... \s ... \s (3)
\s*
$P 10 \p \C :$ \s $2 = 1 + 1$ \s (4)
\s
... \s ... \s (4)
\s*
$(4) \p (3) \p (2, 1+1) [a, b] (P 5) : \C :$ \s $2 \in N$ \s (Theor.)
\end{paracol}

\begin{translateTwoCol}
\vspace{1em}
\emph{Nota.} - Huius facillimae demonstrationis gradus omnes ecplicite scripsimus. Brevitatis causa ipsam ita scribemus:
\vspace{1em}
\ENG
\vspace{1em}
\emph{Note.} - We have explicitly written every step of this very easy proof. For the sake of brevity, we shall write it as follows:
\vspace{1em}
\LAT
$\hspace{1.4cm} P 1 \p 1 [a] (P 6) : \C : 1 +1 \in \N \p P 10 \p (2, 1+1) [a, b] (P 5) : \C : Th$
\ENG
...
\LAT
vel
\ENG
...
\LAT
$\hspace{1.4cm} P 1 \p P 6 : \C : 1 + 1 \in \N \p P 10 \p P 5 : \C : Th$
\ENG
...
\LAT
12. \hspace{0.67cm} $3, 4,... \in N$
\ENG
...
\LAT
13. \hspace{0.67cm} $a,b,c,d \in \N \p a = b \p b = c \p c = d : \C : a = d$
\ENG
...
\LAT
\emph{Dem.} \hspace{0.27cm} $Hyp. P 4 : \C : a,c,d \in \N \p a = c \p c = d \p P4 : \C : Thes.$
\ENG
...
\LAT
14. \hspace{0.67cm} $a,b,c \in \N \p a = b \p b = c \p a \no = c : = \abs$
\ENG
...
\LAT
\emph{Dem.} \hspace{0.27cm} $P4 \p L 39 : \C \p Theor$.
\ENG
...
\LAT
15. \hspace{0.67cm} $a,b,c \in \N \p a = b \p b \no = c : \C \p a \no = c$
\ENG
...
\LAT
16. \hspace{0.67cm} $a,b \in \N \p a + 1 = b + 1 : \C \p a = b$
\ENG
...
\LAT
17. \hspace{0.67cm} $a,b \in \N \p \C : a \no = b \p = \p a + 1 \no = b + 1$
\ENG
...
\LAT
\emph{Dem.} \hspace{0.27cm} $P7 \p L21 : \C \p Theor$
\ENG
...
\end{translateTwoCol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Definition}
\peanoHeadingSmall{Definitio.}
\ENG
\peanoHeadingSmall{Definition.}
\end{translateTwoCol}

\begin{translateTwoCol}
18. \hspace{0.67cm} $a,b \in \N \p \C \p a + (b + 1) = (a + b) + 1$
\ENG
...
\LAT
\vspace{1em}
\emph{Nota.} - Hanc definitionem ita legere oportet: si $a$ et $b$ sunt numeri, et $(a + b) + 1$ sensum habet (scilicet si $a + b$ est numerus), sed $a + (b + 1)$ nondum definitus est, tunc $a + (b + 1)$ significat numerum qui $a + b$ sequitur.
\vspace{1em}
\ENG
\vspace{1em}
\emph{Note.} - This definition should be read: if $a$ and $b$ are numbers, and $(a + b) + 1$ has meaning (that is, if $a + b$ is a number), but $a + (b + 1)$ has not yet been defined, then $a + (b + 1)$ indicates the number that follows $a + b$.
\vspace{1em}
\LAT
Ab hac definitione, et a praecedentibus deducitur:
\ENG
From this definition, and the preceding, we deduce that:
\LAT
\hspace{1.4cm} $a \in \N \p \C \ppp a + 2 = a + (1 + 1) = (a + 1) + 1$
\ENG
...
\LAT
\hspace{1.4cm} $a \in \N \p \C \ppp a + 3 = a + (2 + 1) = (a + 2) + 1$, etc.
\end{translateTwoCol}

\peanoPage{3} % page-number 3

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Theorems}
\peanoHeadingSmall{Theoremata.}
\ENG
\peanoHeadingSmall{Theorems.}
\end{translateTwoCol}

\columnratio{0.08, 0.39, 0.02, 0.08, 0.39, 0.02}
\begin{paracol}{6}
\raggedright
19. \s $a, b \in \N \p \C \p a + b \in N$. \s
\s
... \s ... \s ...
\s*
\emph{Dem.} \s $a \in \N \p$ P6 $: \C : a + 1 \in \N : \C : 1 \in [b \in]$ Ts. \s (1)
\s
... \s ... \s ...
\s*
\s $a \in \N \p \C :: b \in \N \p b \in [b \in]$ Ts $: \C : a + b \in \N \p$ P6 $: \C : (a + b) + 1 \in \N \p$ P18 $: \C : a + (b + 1) \in \N : \C : (b + 1) \in [b \in]$ Ts. \s (2)
\s
... \s ... \s ...
\s*
\s $a \in \N \p (1) \p (2) \p \C : : 1 \in [b \in]$ Ts $\ppp b \in \N \p b \in [b \in]$ Ts $: \C : b + 1 \in [b \in]$ Ts $\ppp ([b \in]$ Ts $) [k]$ P9 $: : \C : \N \C [b \in]$ Ts $\p$ (L50) $:: \C : b \in \N \p C$ Ts. \s (3)
\s
... \s ... \s ...
\s*
\s $(3) \p$ (L42) $: C : a,b \in \N \p \C \p$ Thesis \s (Th.)
\s
... \s ... \s ...
\s*
20. \emph{Def.} \s $a + b + c = (a + b) + c$ \s
\s
... \s ... \s ...
\s*
21. \s $a,b,c \in \N \p \C \p a + b + c \in N$ \s
\s
... \s ... \s ...
\s*
22. \s $a,b,c \in \N \p \C : a = b \p = \p a + c = b + c$ \s
\s
... \s ... \s ...
\s*
\emph{Dem.} \s $a,b \in \N \p$ P7 $: \C \p 1 \in [c \in]$ Ts. \s (1)
\s
... \s ... \s ...
\s*
\s $a,b \in \N \p \C :: c \in \N \p c \in [c \in]$ Ts $\ppp \C \ppp a = b \p = \p a + c = b + c : a + c , b + c \in \N : a + c = b + c \p = \p a + c + 1 = b + c + 1 \ppp \C \ppp a = b \p = \p a + (c + 1) = b + (c + 1) \ppp \C \ppp (c + 1) \in [C\in]$ Ts. \s (2)
\s
... \s ... \s ...
\s*
\s $a,b \in \N \p (1) \p (2) : \C :: 1 \in [c \in]$ Ts $\ppp c \in [c \in]$ Ts $ \p \C \p (c + 1) \in [c \in]$ Ts $:: \C :: c \in \N \p \C \p$ Ts. \s (3)
\s
... \s ... \s ...
\s*
23. \s $a,b,c \in \N \p \C \p a + (b + c) = a + b + c$ \s
\s
... \s ... \s ...
\s*
\emph{Dem.} \s $a,b \in \N \p$ P18 $\p$ P20 $: \C \p 1 \in [c \in]$ Ts. \s (1)
\s
... \s ... \s ...
\s*
\s $a,b \in \N \p \C \ppp c \in \N \p c \in [c \in]$ Ts $: \C : a + (b +c) = a + b + c \p$ P7 $: \C : a + (b + c) + 1 = a + b + c + 1 \p$ P18 $: \C : a + (b + (c + 1)) = a + b + (c + 1) : \C \p c + 1 \in [c \in]$ Ts. \s (2)
\s
... \s ... \s ...
\s*
\s $(1)(2)$ (P9) $\p \C \p$ Theor. \s
\s
... \s ... \s ...
\s*
24. \s $a \in \N \p \C \p 1 + a = a + 1$ \s
\s
... \s ... \s ...
\s*
\emph{Dem.} \s P2 $\p \C \p 1 \in [a \in]$ Ts. \s (1)
\s
... \s ... \s ...
\s*
\s $a \in \N \p a \in [a \in]$ Ts $: \C : 1 + 1 = a + 1 : \C : 1 + (a + 1) = (a + 1) + 1 : \C : (a + 1) \in [a \in]$ Ts. \s (2)
\s
... \s ... \s ...
\s*
\s $(1)(2) \p \C \p$ Theor. \s
\s
... \s ... \s ...
\s*
24'. \s $a,b \in \N \p \C \p 1 + a + b = a + 1 + b$ \s
\s
... \s ... \s ...
\s*
\emph{Dem.} \s Hyp. P24 $: \C \C 1 + a = a + 1 \p$ P22 $: \C \p$ Thesis. \s
\s
... \s ... \s ...
\end{paracol}

\peanoPage{4} % page-number 4

\columnratio{0.08, 0.39, 0.02, 0.08, 0.39, 0.02}
\begin{paracol}{6}
25. \s $a,b\in \N \p \C \p a + b = b + a$. \s
\s
... \s ... \s ...
\s*
\emph{Dem.} \s $a \in \N \p$ P24 $: \C : 1 \in [b \in]$ Ts. \s (1)
\s
... \s ... \s ...
\s*
\s $a \in \N \p \C \ppp b \in \N \p b \in [b \in]$ Ts $: \C : a + b = b + a \p$ P7 $: \C : (a + b) + 1 = (b + a) + 1 \p (a + b) + 1 = (a + (b + 1) \p (b + a ) + 1 = 1 + (b + a) \p 1 + (b + a) = (1 + b) + a \p (1 + b) + a = (b + 1) + a : \C : a + (b + 1) = (b + 1) + a : \C : (b + 1) \in [b \in]$ Ts. \s (2)
\s
... \s ... \s ...
\s*
\s $(1)(2) \p \C \p$ Theor. \s
\s
... \s ... \s ...
\s*
26. \s $a,b,c \in \N \p \C : a = b \p = \p c + a = c + b$. \s
\s
... \s ... \s ...
\s*
27. \s $a,b,c \in \N \p \C : a + b + c = a + c + b$. \s
\s
... \s ... \s ...
\s*
28. \s $a,b,c,d \in \N \p a = b \p c = d : \C \p a + c = b + d$. \s
\s
... \s ... \s ...
\end{paracol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{section}{\S2. Substraction}
\peanoHeadingMedium{\S2. De substractione.}
\ENG
\peanoHeadingMedium{\S2. Substraction.}
\LAT
\phantomsection
\addcontentsline{toc}{subsection}{Explanations}
\peanoHeadingSmall{Explicationes.}
\ENG
\peanoHeadingSmall{Explanations.}
\end{translateTwoCol}

\columnratio{0.1, 0.1, 0.1, 0.2, 0.1, 0.1, 0.1, 0.2}
\begin{paracol}{8}
\centering Signo \s $-$ \s legitur \s \raggedright \emph{minus}.
\s
\centering The sign \s $-$ \s is read \s \raggedright \emph{minus}.
\s*
\centering $\dittoMarkLatin$ \s $<$ \s $\dittoMarkLatin$ \s \raggedright \emph{est minor.}
\s
\centering \dittoMarkEnglish \s $<$ \s \dittoMarkEnglish \s \raggedright \emph{is less than.}
\s*
\centering $\dittoMarkLatin$ \s $>$ \s $\dittoMarkLatin$ \s \raggedright \emph{est maior.}
\s
\centering \dittoMarkEnglish \s $>$ \s \dittoMarkEnglish \s \raggedright \emph{is greater than.}
\end{paracol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Definitions}
\peanoHeadingSmall{Definitiones.}
\ENG
\peanoHeadingSmall{Definitions.}
\end{translateTwoCol}

\columnratio{0.08, 0.39, 0.02, 0.08, 0.39, 0.02}
\begin{paracol}{6}
\raggedright
1. \s $a,b \in \N \p \C : b - a = \N [x \in] (x + a = b)$. \s
\s
\s \s
\s*
2. \s $a,b \in \N \p \C : a < b \p = \p b - a \no = \abs$. \s
\s
\s \s
\s*
3. \s $a,b \in \N \p \C : b > a \p = \p a < b$. \s
\s
\s \s
\s*
\s $a + b - c = (a + b) - c; a - b + c = (a - b) + c; a - b - c = (a - b) - c$. \s
\s \s
\end{paracol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Theorems}
\peanoHeadingSmall{Theoremata.}
\ENG
\peanoHeadingSmall{Theorems.}
\end{translateTwoCol}

\columnratio{0.08, 0.39, 0.02, 0.08, 0.39, 0.02}
\begin{paracol}{6}
\raggedright
4. \s $a,b,a',b' \in \N \p a = a' \p b = b' : \C : b - a = b' - a'$. \s
\s
\s \s
\s*
\emph{Dem.} \s Hyp $\p \C : x + a = b \p = \p x + a' = b' : \C \p$ Thesis. \s
\s
\s \s
\s*
5. \s $a,b \in \N \p \C : a < b \p = \p b - a \in N$. \s
\s
\s \s
\s*
\emph{Dem.} \s $a,b \in \N : \C \ppp x,y \in b - a \p \C_{x,y} : x, y \in \N \p x + a = b \p y + a = b \p$ \S1 P22 $: \C : x = y$. \s (1)
\s
\s \s
\s*
\s $a,b \in \N \p a < b \p$ P2 $\p (1) : \C \ppp b - a \no = \abs : x,y \in b - a \p \C \p x = y : (N, b - a) [s,k]$ (L56) $\ppp \C \ppp b - a \in N$. \s (2)
\s \s
\end{paracol}

\peanoPage{5} % page-number 5

\columnratio{0.08, 0.39, 0.02, 0.08, 0.39, 0.02}
\begin{paracol}{6}
\raggedright
\s $a,b \in \N \p b - a \in \N \p$ (L56) $: \C : b - a \no = \abs : \C : a < b$. \s (3)
\s
\s \s
\s*
\s $(2)(3) \p \C \p$ Theor. \s
\s
\s \s
\s*
6. \s $a,b \in \N \p a < b : \C \p b - a + a = b$. \s
\s
\s \s
\s*
\emph{Dem.} \s Hyp $\p$ P5 $\p$ P1 $: \C : b - a \in \N \p (b - a) \in [x \in] (x + a = b) : \C :$ Thes. \s
\s
\s \s
\s*
7. \s $a,b,c \in \N \p \C : c = b - a \p = \p c + a = b$. \s
\s
\s \s
\s*
\emph{Dem.} \s Hyp $\p \S 1$ P22 $\p$ P6 $: \C : c = b - a \p = \p c + a = b - a + a \p = \p c + a = b$. \s
\s
\s \s
\s*
8. \s $a,b \in \N \p \C \p a + b - a = b$. \s
\s
\s \s
\s*
\emph{Dem.} \s $(a + b, b)[b, c]$ P7 $\p \C \p$ Theor. \s
\s
\s \s
\s*
9. \s $a,b,c \in \N \p a < b : \C : c + (b - a) = c + b - a$. \s
\s
\s \s
\s*
\emph{Dem.} \s Hyp $\p$ P6 $: \C : (b - a) + a = b : \C : c + (b - a) + a = c + b \p$ P7 $: \C :$ Thesis. \s
\s
\s \s
\s*
10. \s $a,b,c \in \N \p a > b + c : \C \p a - (b + c) = a - b - c$. \s
\s
\s \s
\s*
11. \s $a,b,c \in \N \p b > c \p a > b - c : \C \p a - (b - c) = a + c - b$. \s
\s
\s \s
\s*
12. \s $a,b,a',b' \in \N \p a = a' \p b = b' : \C : a < b \p = \p a' < b'$. \s
\s
\s \s
\s*
\emph{Dem.} \s Hyp $\p \C \p b - a = b' - a' \p \C \p b - a \in \N = b' - a' \in \N \p \C \p$ Thes. \s
\s
\s \s
\s*
13. \s $a,b \in \N \p \C \p a < a + b$. \s
\s
\s \s
\s*
\emph{Dem.} \s Hyp $\p P8 : \C : a + b - a = b : \C \p a + b - a \in \N \p$ P5 $: \C :$ Thesis. \s
\s
\s \s
\s*
14. \s $a,b,c \in \N \p a < b \p b < c : \C \p a < c$. \s
\s
\s \s
\s*
\emph{Dem.} \s Hyp $\p \C : b - a \in \N \p c - b \in : \C : (b - a) + (c - b) \in \N : \C : c - a \in \N : \C \p$ Thesis. \s
\s
\s \s
\s*
15. \s $a,b,c \in \N \p \C : a < b \p = \p a + c < b + c$. \s
\s
\s \s
\s*
\emph{Dem.} \s Hyp $\p \C : a < b \p = \p b - a \in \N \p = \p (b + c) - (a + c) \in \N \p = \p a + c < b + c$. \s
\s
\s \s
\s*
16. \s $a,b,a',b' \in \N \p a < b \p a' < b' : \C \p a + a' < b + b'$. \s
\s
\s \s
\s*
\emph{Dem.} \s Hyp $\p \C : a + a' < b + a' \p b + a' < b + b' : \C \p$ Thesis. \s
\s
\s \s
\s*
17. \s $a,b,c \in \N \p a < b < c : \C \p c - a > c - b$. \s
\s
\s \s
\s*
\emph{Dem.} \s Hyp $\p \C b - a \in \N \p c - b \in \N \p (c - b) + (b - a) = c - a : \C \p$ Thesis. \s
\s
\s \s
\s*
18. \s $a \in \N \p \C : a = 1 \p \cup \p a > 1$. \s
\s
\s \s
\s*
\emph{Dem.} \s $1 \in [a \in]$ Thesis. \s
\s
\s \s
\s*
\s $a \in \N \p$ P13 $: \C : a + 1 > 1 : \C : a + 1 \in [a \in]$ Thesis. \s
\s
\s \s
\s*
\s $(1)(2) \p \C \p$ Theor. \s
\s
\s \s
\end{paracol}

\peanoPage{6} % page-number 6

\columnratio{0.08, 0.39, 0.02, 0.08, 0.39, 0.02}
\begin{paracol}{6}
\raggedright
19. \s $a,b \in \N \p \C \p a + b \no = b$.\s
\s
\s \s
\s*
\emph{Dem.} \s $a \in \N \p \S1$ P8 $: \C : a + 1 \no = 1 : \C : 1 \in [b \in]$ Thesis. \s (1)
\s
\s \s
\s*
\s $a \in \N \p b \in \N \p b \in [b \in]$ Ts $: \C : a + b \no = b \p \S1$ P17 $: \C : a + (b + 1) \no = b + 1 : \C : b + 1 \in [b \in]$ Ts. \s (2)
\s
\s \s
\s*
\s $(1)(2) \p \C \p$ Theor. \s
\s
\s \s
\s*
20. \s $a,b \in \N \p a < b \p a = b : = \abs$. \s
\s
\s \s
\s*
\emph{Dem.} \s Hyp $: \C : b - a \in \N \p (b - a) + a = a \p$ P19 $: \C : \abs$. \s
\s
\s \s
\s*
21. \s $a,b \in \N \p a > b \p a = b : = \abs$. \s
\s
\s \s
\s*
22. \s $a,b \in \N \p a > b \p a < b : = \abs$. \s
\s
\s \s
\s*
23. \s $a,b \in \N : \C : a < b \p \cup \p a = b \p \cup \p a > b$. \s
\s
\s \s
\s*
\emph{Dem} \s $a \in \N \p$ P18 $: \C \p 1 \in [b \in]$ Ts. \s (1)
\s
\s \s
\s*
\s $a,b \in \N \p a < b : \C \p a < b + 1$. \s (2)
\s
\s \s
\s*
\s $a,b \in \N \p a = b : \C \p a < b + 1$. \s (3)
\s
\s \s
\s*
\s $a,b \in \N \p a > b : \C : a - b \in \N \p$ P18 $: \C : a - b = 1 \p \cup \p a - b > 1$. \s (4)
\s
\s \s
\s*
\s $a,b \in \N \p a - b = 1 : \C \p a = b + 1$. \s (5)
\s
\s \s
\s*
\s $a,b \in \N \p a - b > 1 : \C \p a > b + 1$. \s (6)
\s
\s \s
\s*
\s $a,b \in \N \p a > b \p (4)(5)(6) : \C : a = b + 1 \p \cup \p a > b + 1 \p$ \s (7)
\s
\s \s
\s*
\s $a,b \in \N : a > b \p \cup \p a = b \p \cup \p a > b : (2)(3)(7) \ppp \C \ppp a < b + 1 \p \cup \p a = b + 1 \p \cup \p a > b + 1$. \s (8)
\s
\s \s
\s*
\s $a,b \in \N \p b \in [b \in]$ Ts $\p (8) : \C : b + 1 \in [b \in]$ Ts. \s (9)
\s
\s \s
\s*
\s $(1)(9) \p \C \p$ Theor. \s
\s
\s \s
\end{paracol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{section}{\S3. Maxima and minima}
\peanoHeadingMedium{\S3. De maximis et minimis.}
\s
\peanoHeadingMedium{\S3. Maxima and minima.}
\s*
\phantomsection
\addcontentsline{toc}{subsection}{Explanations}
\peanoHeadingSmall{Explicationes.}
\s
\peanoHeadingSmall{Explanations.}
\s*
Sit $a \in K N$, hoc est sit $a$ quaedam numerorum classis; tunc $M a$ legatur \emph{maximus inter} $a$, et $\mini a$ legatur \emph{minimus inter a}.
\s
Let $a \in \K \N$, that is, let $a$ be a class of numbers; then $\M a$ is read \emph{greatest among} $a$, and $\mini a$ is read \emph{least among a}.
\s*
\phantomsection
\addcontentsline{toc}{subsection}{Definitions}
\peanoHeadingSmall{Definitiones.}
\s
\peanoHeadingSmall{Definitions.}
\end{translateTwoCol}

\columnratio{0.08, 0.39, 0.02, 0.08, 0.39, 0.02}
\begin{paracol}{6}
\raggedright
1. \s $a \in K \N \p \C : M a = [x \in] (x \in a \ppp a \p \in \larger x : = \abs$.\s
\s
\s \s
\s*
2. \s $a \in K \N \p \C : \mini a = [x \in] (x \in a \ppp a \p \in \smaller x : = \abs$.\s
\end{paracol}

\peanoPage{7} % page-number 7

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Theorems}
\peanoHeadingSmall{Theoremata.}
\s
\peanoHeadingSmall{Theorems.}
\end{translateTwoCol}

\columnratio{0.08, 0.39, 0.02, 0.08, 0.39, 0.02}
\begin{paracol}{6}
\raggedright
3. \s $n \in \N \p a \in K \N \p a \no = \abs \p a \such \larger n = \abs : \C \p M a \in N$.\s
\s
\s \s
\s*
\emph{Dem}. \s $a \in K \N \p a \no = \abs \p a \such \larger 1 = \abs : \C : a = 1 : \C \p M a = 1 : \C \p M a \in N$.\s (1)
\s
\s \s
\s*
\s $(1) \C : 1 \in [n \in]$ (Hp $\C$ Ts). \s (2)
\s
\s \s
\s*
\s $n \in \N \p a \in K \N \p a \such \larger n + 1 = \abs \p n + 1 \in a : \C : n + 1 = M a : \C : M a \in N$. \s (3)
\s
\s \s
\s*
\s $n \in \N \p a \in K \N \p a \such \larger n + 1 = \abs \p n + 1 \no \in a : \C : a \such \larger n = \abs$. \s (4)
\s
\s \s
\s*
\s $n \in [n \in]$ (Hp $\C$ Ts) $\p a \in K \N \p a \such > n + 1 = \abs \p n + 1 \no \in a : \C : M a \in N$. \s (5)
\s
\s \s
\s*
\s $n \in [n \in]$ (Hp $\C$ Ts) $. (6) : \C \p (n+ 1) \in [n \in]$ (Hp $\C$ Ts). \s (7)
\s
\s \s
\s*
\s $(2)(7) \p \S1$ P9 $: \C : n \in \N \p \C \p$ Hp $\C$ Ts. \s (Th.)
\s
\s \s
\s*
4. \s $a \in K \N \p a \no = \abs : \C \p \mini a \in N$. \s
\s
\s \s
\s*
5. \s $a \in K \N \p \C : \mini a = M [x \in] (a \such \smaller x = \abs)$. \s
\s
\s \s
\end{paracol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{section}{\S4. Multiplication}
\peanoHeadingMedium{\S4. De multiplicatione.}
\s
\peanoHeadingMedium{\S4. Multiplication.}
\s*
\phantomsection
\addcontentsline{toc}{subsection}{Definitions}
\peanoHeadingSmall{Definitiones.}
\s
\peanoHeadingSmall{Definitions.}
\end{translateTwoCol}

\columnratio{0.08, 0.39, 0.02, 0.08, 0.39, 0.02}
\begin{paracol}{6}
\raggedright
1. \s $a \in \N \p \C \p a \times 1 = a$.\s
\s
\s \s
\s*
2. \s $a,b \in \N \p \C \p a \times (b + 1) = a \times b + a$. \s
\s
\s \s
\s*
\s $a b = a \times b; a b + c = (a b) + c; a b c = (a b) c$. \s
\s
...
\end{paracol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Theorems}
\peanoHeadingSmall{Theoremata.}
\s
\peanoHeadingSmall{Theorems.}
\end{translateTwoCol}

\columnratio{0.08, 0.39, 0.02, 0.08, 0.39, 0.02}
\begin{paracol}{6}
\raggedright
3. \s $a,b \in \N \p \C \p ab \in N$.\s
\s
\s \s
\s*
\emph{Dem}. \s $a,b \in \N  \p$ P1 $: \C : a \times 1 \in \N : \C \p 1 \in [b \in]$ Ts. \s
\s
\s \s
\s*
\s $a,b \in \N \p b \in [b \in]$ Ts $: \C : a \times b \in \N \p \S 1$ P19 $: \C : ab + a \in \N \p$ P1 $: \C : a (b + 1) \in \N : \C : b + 1 \in [b \in]$ Ts. \s (2)
\s
\s \s
\s*
\s $(1)(2) \p \C \p$ Theor. \s
\s
\s \s
\end{paracol}

\peanoPage{8} % page-number 8

\columnratio{0.1, 0.3, 0.1, 0.1, 0.3, 0.1}
\begin{paracol}{6}
\raggedright
4. \s $a,b,c \in \N \p \C \p (a + b) c = a c  b c$.\s
\s
\s \s
\s*
\emph{Nota.} \s Haec est prop. $5^a$ Euclidis elem. libri VII. \s
\s
\emph{Note.} \s This is prop. $5^a$ of Euclid's elem. book VII. \s
\s*
\emph{Dem.} \s $a,b \in \N \p$ P1 $: \C : 1 \in [c \in]$ Ts. \s (1)
\s
\s \s
\s*
\s $a,b,c \in \N \p c \in [c \in]$ Ts $: \C : (a + b) c = a c + b c \p \S1$ P22 $: \C : (a + b) c + a + b = ac + bc + a + b \p$ P2 $: \C : (a + b) (c + 1) = a (c + 1) + b (c + 1) : \C : c + 1 \in [c \in]$ Ts. \s (2)
\s
\s \s
\s*
\s $(1)(2) \p \C \p$ Theor. \s
\s
\s \s
\s*
5. \s $a \in \N \p \C \p 1 \times a = a \p$ \s
\s
\s \s
\s*
\emph{Dem.} \s $1 \in [a \in]$ Ts. \s (1)
\s
\s \s
\s*
\s $a \in [a \in]$ Ts $\p \C \p 1 \times a = a \p \C \p 1 \times a + 1 = a + 1 \p \times \p 1 \times (a + 1) = a + 1 \p \C \p a + 1 \in [a \in]$ Ts. \s (2)
\s
\s \s
\s*
\s $(1)(2) \p \C \p$ Theor. \s
\s
\s \s
\s*
6. \s $a,b \in \N \p \C \p b a + a = (b + 1) a$. \s
\s
\s \s
\s*
7. \s $a,b \in \N \p \C \p ab = ba$. \s (Eucl. VII, 16)
\s
\s \s
\s*
\emph{Dem.} \s $a \in \N \p$ P5 $\p$ P1 $: \C \p a \times 1 = a = 1 \times a : \C : 1 \in [b \in]$ Ts. \s (1)
\s
\s \s
\s*
\s $a,b \in \N \p b \in [b \in]$ Ts $: \C : ab = ba : \C : ab + a = ba + a \p$ P1 $\p $P6$ : \C : a (b + 1) = (b + 1) a : \C : b + 1 \in [b \in]$ Ts. \s (2)
\s
\s \s
\s*
\s $(1)(2) \p \C \p$ Theor. \s
\s
\s \s
\s*
8. \s $a,b,c \in \N \p \C \p a (b + c) = ab + ac$. \s
\s
\s \s
\s*
\emph{Dem.} \s P4 $\p$ P7 $: \C \p$ Theor. \s
\s
\s \s
\s*
9. \s $a,b,c \in \N \p a = b : \C : ac = bc$. \s
\s
\s \s
\s*
\emph{Dem.} \s $a,b \in \N \p a = b :: \C :: 1 \in [c \in]$ Ts $\ppp c \in [c \in]$ Ts $\p \C : ac = bc \p a = b : \C : ac + a = bc + b : \C : a (c + 1) = b (c + 1) : \C : c + 1 \in [c \in]$ Ts $:: \C : c \in \N \p \C \p$ Ts. \s
\s
\s \s
\s*
10. \s $a,b,c \in \N : a \smaller b : \C \p (b-a)c = bc - ac$. \s (Eucl. VII, 7)
\s
\s \s
\s*
\emph{Dem.} \s Hyp $\p \C : b - a \in \N \p (b-a) + a = b : \C : (b-a)c + ac=bc : \C : (b-a)c = bc-ac$. \s
\s
\s \s
\s*
11. \s $a,b,c \in \N \p a \smaller b : \C : ac \smaller bc$. \s
\s
\s \s
\s*
\emph{Dem.} \s Hyp $\p \C : b - a \in \N \p$ P3 $: \C : (b-a) c \in \N \p$ P10 $: \C : bc - ac \in \N : \C$ Thesis. \s
\s
\s \s
\s*
12. \s $a,b,c \in \N \p \C \ppp a \smaller b \p = \p ac \smaller bc : a = b \p = \p ac = bc : a \larger b \p = \p ac \larger bc$. \s
\s
\s \s
\s*
13. \s $a,b,a',b' \in \N \p a \smaller a' \p b \smaller b' : \C : ab \smaller a'b'$. \s
\s
\s \s
\s*
14. \s $a,b \in \N : \C \p ab \p \larger \cup = \p a$. \s
\s
\s \s
\s*
15. \s $a,b,c \in \N \p \C \p a(bc)=abc$. \s
\s
\s \s
\end{paracol}

\peanoPage{9} % page-number 9

\columnratio{0.1, 0.3, 0.1, 0.1, 0.3, 0.1}
\begin{paracol}{6}
\raggedright
\emph{Dem.} \s $a,b \in \N \p$ P1 $: \C : 1 \in [c \in]$ Ts. \s (1)
\s
\s \s
\s*
\s $a,b,c \in \N \p c \in [c \in]$ Ts $: \C : a(bc) = abc : \C : a(bc) + ab = abc + ab : \C : a (bc + b) = ab (c+ 1) : \C : a (b (c +1)) = ab (c + 1) : \C : c + 1 \in [c \in]$ Ts. \s (2)
\s
\s \s
\s*
\s $(1)(2) \p \C \p$ Theor. \s
\s
\s \s
\end{paracol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{section}{\S5. Powers}
\peanoHeadingMedium{\S5. De potestatibus.}
\s
\peanoHeadingMedium{\S5. Powers.}
\s*
\phantomsection
\addcontentsline{toc}{subsection}{Definitions}
\peanoHeadingSmall{Definitiones.}
\s
\peanoHeadingSmall{Definitions.}
\end{translateTwoCol}

\columnratio{0.1, 0.3, 0.1, 0.1, 0.3, 0.1}
\begin{paracol}{6}
\raggedright
1. \s $a \in \N \p \C \p a^1 = a$. \s
\s
\s \s
\s*
2. \s $a,b \in \N \p \C \p a^{b+1} = a^{b} a$. \s
\s
\s \s
\end{paracol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Theorems}
\peanoHeadingSmall{Theoremata.}
\s
\peanoHeadingSmall{Theorems.}
\end{translateTwoCol}

\columnratio{0.1, 0.3, 0.1, 0.1, 0.3, 0.1}
\begin{paracol}{6}
\raggedright
3. \s $a,b \in \N \p \C \p a^{b} \in N$. \s
\s
\s \s
\s*
\emph{Dem.} \s $a \in \N \p$ P1 $: \C \p 1 \in [b \in]$ Ts. \s (1)
\s
\s \s
\s*
\s $a,b \in \N \p b \in [b \in]$ Ts $: \C : a^b \in \N \p \S4 $P3$ : \C : a^b a \in \N \p$ P1$: \C : a^{b+1} \in \N : \C : b + 1 \in [b \in]$ Ts. \s (2)
\s
\s \s
\s*
\s $(1)(2) \p \C \p.$ Theor. \s
\s
\s \s
\s*
4. \s $a \in \N \p \C \p 1^{a} = 1$. \s
\s
\s \s
\s*
5. \s $a,b,c \in \N \p \C \p a^{b + c} = a^b a^c$. \s
\s
\s \s
\s*
6. \s $a,b,c \in \N \p \C \p (ab)^c = a^c b^c$. \s
\s
\s \s
\s*
7. \s $a,b,c \in \N \p \C \p (a^b)^c = a^{bc}$. \s
\s
\s \s
\s*
8. \s $a,b,c \in \N \p \C \ppp a \smaller b \p = \p a^c \smaller b^c : a = b \p = \p a^c = b^c : a \larger b \p = \p a^c \larger b^c$. \s
\s
\s \s
\s*
9. \s $a,b,c \in \N \p a \larger 1 \p \C \ppp b \smaller c \p = \p a^b \smaller a^c : b = c \p = \p a^b = a^c : b \larger c \p = \p a^b \larger a^c$.
\end{paracol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{section}{\S6. Division}
\peanoHeadingMedium{\S6. De divisione.}
\s
\peanoHeadingMedium{\S6. Division.}
\s*
\phantomsection
\addcontentsline{toc}{subsection}{Explications}
\peanoHeadingSmall{Explicationes.}
\s
\peanoHeadingSmall{Explications.}
\end{translateTwoCol}

\columnratio{0.12, 0.06, 0.06, 0.26, 0.12, 0.06, 0.06, 0.26}
\begin{paracol}{8}
\centering Signum \s \raggedright $/$ \s \centering legatur \s \raggedright \emph{divisus per.}
\s
\centering The sign \s \raggedright $/$ \s \centering is read \s \raggedright \emph{divided by.}
\s*
\centering $\dittoMarkLatin$ \s \raggedright $\D$ \s \centering $\dittoMarkLatin$ \s \raggedright \emph{dividit}, sive \emph{est divisor.}
\s
\centering \dittoMarkEnglish \s \raggedright $\D$ \s \centering \dittoMarkEnglish \s \raggedright \emph{divides}, or \emph{is a divisor of.}
\s*
\centering $\dittoMarkLatin$ \s \raggedright $\mult$ \s \centering $\dittoMarkLatin$ \s \raggedright \emph{est multiplex.}
\s
\centering \dittoMarkEnglish \s \raggedright $\mult$ \s \centering \dittoMarkEnglish \s \raggedright \emph{is a multiple of.}
\s*
\centering $\dittoMarkLatin$ \s \raggedright $\Np$ \s \centering $\dittoMarkLatin$ \s \raggedright \emph{numerus primus.}
\s
\centering \dittoMarkEnglish \s \raggedright $\Np$ \s \centering \dittoMarkEnglish \s \raggedright \emph{prime number.}
\s*
\centering $\dittoMarkLatin$ \s \raggedright $\otherpi$ \s \centering $\dittoMarkLatin$ \s \raggedright \emph{est primus cum.}
\s
\centering \dittoMarkEnglish \s \raggedright $\otherpi$ \s \centering \dittoMarkEnglish \s \raggedright \emph{is prime with.}
\end{paracol}

\peanoPage{10} % page-number 10

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Definitions}
\peanoHeadingSmall{Definitiones.}
\s
\peanoHeadingSmall{Definitions.}
\end{translateTwoCol}

\columnratio{0.1, 0.3, 0.1, 0.1, 0.3, 0.1}
\begin{paracol}{6}
\raggedright
1. \s $a,b \in \N \p \C : b / a = \N [x \in] (xa=b)$. \s
\s
\s \s
\s*
2. \s $a,b \in \N \p \C : a D b \p = \p b / a \no = \abs$. \s
\s
\s \s
\s*
3. \s $a,b \in \N \p \C : b \mult a \p = \p a D b$. \s
\s
\s \s
\s*
4. \s $\Np = \N [x \in] (\such D x \p \such \larger 1 \p \such \smaller x : = \abs)$. \s
\s
\s \s
\s*
5. \s $a,b \in \N \p \C :: a \pi b \ppp = \ppp \such D a \p \such D b \p \such \larger 1 : = \abs$. \s
\s
\s \s
\s*
6. \s $a,b \in \N \p \C \ppp \such D (a,b) : = : \such D a \p \cap \p \such D b$. \s
\s
\s \s
\s*
7. \s $a,b \in \N \p \C \ppp \such \mult (a,b) : = : \such \mult a \p \cap \p \such \mult b$. \s
\s
\s \s
\s*
\s $ab / c = (ab)/c; a/b/c = (a/b)/c; a/b \times c = (a/b)c$. \s
\s
\s \s
\end{paracol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Theorems}
\peanoHeadingSmall{Theoremata.}
\s
\peanoHeadingSmall{Theorems.}
\end{translateTwoCol}

\columnratio{0.1, 0.4, 0.1, 0.4}
\begin{paracol}{4}
\raggedright
\emph{Nota.} \s Haec theoremata ut in substractione demonstrantur.
\s
\emph{Note.} \s These theorems are proved as for subtraction.
\end{paracol}

\columnratio{0.1, 0.3, 0.1, 0.1, 0.3, 0.1}
\begin{paracol}{6}
\raggedright
8. \s $a,b,a',b' \in \N \p a = a' \p b = b' : \C \p a / b = a' / b'$. \s
\s
\s \s
\s*
9. \s $a,b,a',b' \in \N \p a = a' \p b = b' : \C : a D b \p = \p a' D b'$. \s
\s
\s \s
\s*
10. \s $a,b,c \in \N \p \C : ac = b \p = \p c = b / a$. \s
\s
\s \s
\s*
11. \s $a,b \in \N \p \C : a D b \p = \p b / a \in N$. \s
\s
\s \s
\s*
12. \s $a \in \N \p \C \p a / 1 = a$. \s
\s
\s \s
\s*
13. \s $a \in \N \p \C \p a / a = 1$. \s
\s
\s \s
\s*
14. \s $a \in \N \p \C \p 1 D a$. \s
\s
\s \s
\s*
15. \s $a \in \N \p \C \p a D a$. \s
\s
\s \s
\s*
16. \s $a,b \in \N \p ab / b = a$. \s
\s
\s \s
\s*
17. \s $a,b \in \N \p a D b : \C \p a (b/a) = b$. \s
\s
\s \s
\s*
18. \s $a,b,c \in \N \p c D b : \C \p a (b/c) = ab / c$. \s
\s
\s \s
\s*
19. \s $a,b,c \in \N \p a \mult bc : \C : a / (bc) = a / b / c$. \s
\s
\s \s
\s*
20. \s $a,b,c \in \N \p a \mult b \p b \mult c : \C \p a / (b/c) = a / b \times c$. \s
\s
\s \s
\s*
21. \s $a,m,n \in \N \p m \larger n : \C \p a^m / a^n = a^{m-n}$. \s
\s
\s \s
\s*
22. \s $a,b \in \N \p \C \p a D ab$. \s
\s
\s \s
\s*
23. \s $a,b,c \in \N \p a D b \p b D c : \C \p a D c$. \s
\s
\s \s
\s*
24. \s $a,b,c \in \N \p a D b D c : \C \p c / a \mult c / b$. \s
\s
\s \s
\s*
25. \s $a,b,c \in \N \p c D a \p c D b : \C \p (a + b) / c = a / c + b / c$. \s
\s
\s \s
\s*
26. \s $a,b,c \in \N \p c D a \p c D b \p a \larger b : \C : (a - b) / c = a / c - b / c$. \s
\s
\s \s
\s*
27. \s $a,b,c, \in \N \p c D a \p c D b : \C \p c D a + b$. \s
\s
\s \s
\s*
28. \s $a,b,c \in \N \p c D a \p c D b \p a \larger b : \C \p c D a - b$. \s
\end{paracol}

\peanoPage{11} % page-number 11

\columnratio{0.1, 0.3, 0.1, 0.1, 0.3, 0.1}
\begin{paracol}{6}
\raggedright
29. \s $a,b,c,m,n \in \N \p c D a \p c D b : \C \p c D ma + nb$. \s
\s
\s \s
\s*
30. \s $a,b,c,m,n \in \N \p c D a \p c D b \p ma \larger nb : \C \p c D ma - nb$. \s
\s
\s \s
\s*
31. \s $a,b \in \N \p a D b : \C : a \p \smaller \cup = \p b$. \s
\s
\s \s
\s*
\emph{Dem.} \s Hyp . P11 $\p$ P17 $\p$ \S4 P14 $: \C : b / a \in \N \p a (b / a) = b \p a \smaller \cup = a (b / a) : \C \p$ Thesis. \s
\s
\s \s
\s*
32. \s $a,b \in \N \p a D b \p b D a : \C \p a = b$. \s
\s
\s \s
\s*
33. \s $a \in \N \p \C \p M \such D a = a$. \s
\s
\s \s
\s*
34. \s $a,b \in \N \p a > b : \C \p \such  D (a,b) = \such D (b,a - b)$. \s
\s
\s \s
\s*
\emph{Dem.} \s Hyp. P28 $: \C \ppp x D a \p x D b : \C : x D b \p x D (a-b)$ \s (1)
\s
\s \s
\s*
\s Hyp. P27 $: \C \ppp x D b \p x D (a-b) : \C : x D b \p x D (b+(a-b)) : \C : x D b \p x D a$. \s (2)
\s
\s \s
\s*
\s $(1)(2) \C :$ Hyp. $\C \ppp x D a \p x D b : = : x D b \p x D(a-b)$. \s (Th.)
\s
\s \s
\s*
35. \s $a,b \in \N \p \C : M \such D (a,b) \in N$. \s
\s
\s \s
\s*
\emph{Dem.} \s $1 D a \p 1 D b : \C : \such D (a,b) \no = \abs$. \s (1)
\s
\s \s
\s*
\s $\such D (a,b) \p \such \larger a : = \abs$. \s (2)
\s
\s \s
\s*
\s $(1)(2) \p \S3$ P3 $: \C \p$ Th. \s
\s
\s \s
\s*
36. \s $a,b \in \N \p \C \p \such D (a,b) = \such D M \such D (a,b).$ \s (Eucl. VII, 2)
\s
\s \s
\s*
\emph{Dem.} \s $k = \N [c \in]$ (Hp. $a \smaller c \p b \smaller c : \C \p$ Ts.). \s (1)
\s
\s \s
\s*
\s $a \in \N \p b \in \N \p a \smaller 1 \p b \smaller 1 : = \abs$. \s (2)
\s
\s \s
\s*
\s $ (1)(2) \p \C \p 1 \in K$. \s (3)
\s
\s \s
\s*
\s $a,b \in \N \p a \smaller c + 1 \p b \smaller c + 1 : \C \ppp a \smaller c \p b \smaller c : \cup : a = c \p b \smaller c : \cup : a \smaller c \p b = c : \cup : a = c \p b = c$. \s (4)
\s
\s \s
\s*
\s $c \in k \p a,b \in \N \p a < c \p b < c : \C :$ Ts. \s (5)
\s
\s \s
\s*
\s $c \in k \p a,b \in \N \p a = c \p b \smaller c : \C : c \in k \p b \smaller c /p a - b \smaller c \p \such D (a,b) = \such D (b,a-b) : \C : \such D (b, a-b) = \such D m \such D (b,a-b) : \C : \such D (a,b) = \such D M \such D (a,b) : \C :$ Ts. \s (6)
\s
\s \s
\s*
\s $(a,b) [b,a] (6) \C \p c \in k \p a,b \in \N \p a \smaller c \p b = c : \C :$ Ts. \s (7)
\s
\s \s
\s*
\s $c \in k \p a,b \in \N \p a = c \p b = c : \C : \such D (a,b) = \such D c = \such D M \such D c = \such D M \such D (a,b) : \C :$ Ts. \s (8)
\s
\s \s
\s*
\s $(4)(5)(6)(7)(8) \p \C \p c \in k \p a,b \in \N \p a \smaller c + 1 \p b \smaller c + 1 : \C :$ Ts. \s (9)
\s
\s \s
\s*
\s $(9) \C \p c \in k \p \C \p (c + 1) \in k$. \s (10)
\s
\s \s
\s*
\s $(1)(10) \p \C \ppp c \in \N \p$ Hp. $a \smaller c \p b \smaller c : \C :$ Ts. \s (11)
\s
\s \s
\s*
\s $(a + b) [c] (11) \p \C :$ Hp. $ \C \p$ Ts. \s (Th.)
\s
\s \s
\s*
37. \s $a,b,m \in \N \p \C \p M \such D (am, bm) = m \times M \such D (a,b)$. \s
\s
\s \s
\end{paracol}

\peanoPage{12} % page-number 12

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{section}{\S7. Various theorems}
\peanoHeadingMedium{\S7. Theoremata varia.}
\s
\peanoHeadingMedium{\S7. Various theorems.}
\end{translateTwoCol}

\columnratio{0.1, 0.3, 0.1, 0.1, 0.3, 0.1}
\begin{paracol}{6}
\raggedright
1. \s $a,b \in \N \p a^2 + b^2 \mult 7 : \C : a \mult 7 \p b \mult 7$. \s
\s
\s \s
\s*
2. \s $x \in \N \p \C \p x (x + 1) \mult 2$. \s
\s
\s \s
\s*
3. \s $x \in \N \p \C \p x (x + 1) (x + 2) \mult 6$. \s
\s
\s \s
\s*
4. \s $x \in \N \p \C \p x (x + 1) (2x + 1) \mult 6$. \s
\s
\s \s
\s*
5. \s $x \in \N \p \C : x \p \pi \p x + 1$. \s
\s
\s \s
\s*
6. \s $x \in \N \p \C : 2x - 1 \p \pi \p 2x + 1$. \s
\s
\s \s
\s*
7. \s $x \in \N \p \C \p (2x + 1)^2 - 1 \mult 8$. \s
\s
\s \s
\s*
8. \s $a \in \N \p a \larger 1 : \C \ppp \Np \p \such \larger 1 \p \such D a : \no = \abs$. \s (Eucl. VII, 31)
\s
\s \s
\s*
9. \s $a,b \in \N \ppp b^2 \larger a \ppp \such D a \p \such \larger 1 \p \such \smaller b : = \abs :: \C \p a \in \Np$. \s
\s
\s \s
\s*
10. \s $a,b \in \N \p a \in \Np \p a \no D b : \C : a \pi b$. \s (Eucl. VII, 29)
\s
\s \s
\s*
11. \s $a,b,c \in \N \p a D bc \p a \pi b : \C \p a D c$. \s
\s
\s \s
\s*
12. \s $a,b \in \N \p m = M \such D (a,b) : \C : a / m \pi \p b / m$. \s
\s
\s \s
\s*
13. \s $ a \in \Np \p b,c \in \N \p a D b c : \C : a D b \p \cup \p a D c$. \s (Eucl. VII, 30)
\s
\s \s
\s*
14. \s $a \in \Np \p b,n \in \N : \C : a D b^n \p = \p a D b$. \s (Eucl. IX, 12)
\s
\s \s
\s*
15. \s $a,b,c \in \N \p a \pi b \p c D a : \C : c \pi b$. \s (Eucl. VII, 23)
\s
\s \s
\s*
16. \s $ a,b,c \in \N \p \C \ppp a \pi b \p a \pi c : = : a \pi bc$. \s (Eucl. VII, 24)
\s
\s \s
\s*
17. \s $a,b,c \in \N \p b \pi c \p b D a \p c D a : \C \p bc D a$. \s
\s
\s \s
\s*
18. \s $a,b,c \in \N \ p a \pi b : \C : \such D (ac, b) = \such D (c,b)$. \s
\s
\s \s
\s*
19. \s $a,b \in \N \p \C \p \mini \such \mult (a,b) \in N$. \s
\s
\s \s
\s*
20. \s $a,b \in \N \p \C \p \mini \such \mult (a,b) = ab / M \such D (a,b)$. \s (Eucl. VII, 34)
\s
\s \s
\s*
21. \s $ a,b,c \in \N \p c \mult a \p c \mult b : \C : c \mult \mini \such D (a,b)$. \s (Eucl. VII, 35)
\s
\s \s
\s*
22. \s $x \in \N \p x < 41 : \C \p 41 - x + x^2 \in \Np$. \s
\s
\s \s
\s*
23. \s $M \p \Np : = \abs$. \s (Eucl. IX, 20)
\s
\s \s
\s*
23. \s $n \in \Np \p a \in \N \p a \no \mult n : \C \p a^{n-1} - 1 \mult n$. \s (Fermat)
\end{paracol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{section}{\S8. Rational numbers}
\peanoHeadingMedium{\S8. Numerorum rationes.}
\s
\peanoHeadingMedium{\S8.Rational numbers.}
\s*
\phantomsection
\addcontentsline{toc}{subsection}{Explications}
\peanoHeadingSmall{Explicationes.}
\s
\peanoHeadingSmall{Explications.}
\end{translateTwoCol}

\begin{translateTwoCol}
\raggedright
Si $p,q \in \N$, tunc $\frac{p}{q}$ legitur \emph{ratio numeri} $p$ \emph{numero} $q$.
\s
If $p,q \in \N$, then $\frac{p}{q}$ is read \emph{the ratio of the number} $p$ {to the number} $q$.
\s*
Signum $\R$ legitur \emph{duorum numerorum ratio}, et indicat numeros rationales positivos.
\s
The sign $\R$ is read \emph{ratio of two numbers}, and indicates the positive rational numbers.
\end{translateTwoCol}

\peanoPage{13} % page-number 13

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Definitions}
\peanoHeadingSmall{Definitiones.}
\s
\peanoHeadingSmall{Definitions.}
\end{translateTwoCol}

\columnratio{0.1, 0.3, 0.1, 0.1, 0.3, 0.1}
\begin{paracol}{6}
\raggedright
1. \s $m,p,q \in \N \p \C \p m \frac{p}{q} = m p / q$. \s
\s
\s \s
\s*
2. \s $p,q,p',q' \in \N \p \C :: \frac{p}{q} = \frac{p'}{q'} \p = \ppp x \in \N \p x \frac{p}{q}, x \frac{p'}{q'} \in \N : \C_x \p x \frac{p}{q} = x \frac{p'}{q'}$. \s
\s
\s \s
\s*
3. \s $R = :: [x \in] \ppp p, q \in \N \p \frac{p}{q} = x : \no = \abs$. \s
\s
\s \s
\s*
4. \s $p \in \N \p \C \p \frac{p}{1} = p$. \s
\end{paracol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Theorems}
\peanoHeadingSmall{Theoremata.}
\s
\peanoHeadingSmall{Theorems.}
\end{translateTwoCol}

\columnratio{0.1, 0.3, 0.1, 0.1, 0.3, 0.1}
\begin{paracol}{6}
\raggedright
5. \s $p,q,p',q' \in \N \p \C :: \frac{p}{q} = \frac{p'}{q'} \p = \p pq' = p'q$. \s (Eucl. VII, 19)
\s
\s \s
\s*
\emph{Dem.} \s Hp. $\frac{p}{q} = \frac{p'}{q'} : \C \ppp qq', qq' \frac{p}{q}, qq' \frac{p'}{q'} \in \N \p$ P2 $\ppp \C \ppp qq' \frac{p}{q} = qq' \frac{p'}{q'} \p qq' \frac{p}{q} = pq' \p qq' \frac{p'}{q'} = p'q \ppp \C \ppp pq' = p'q$. \s (1)
\s
\s \s
\s*
\s Hp. $pq' = p'q \ppp \C \ppp x \in \N \p x \frac{p}{q}, x \frac{p'}{q'} \in \N : \C_x : x p'q' = x p'q : \C : (x \frac{p}{q}) qq' = (x \frac{p'}{q'}) qq' : \C : x \frac{p}{q} = x \frac{p'}{q'}$. \s (2)
\s
\s \s
\s*
\s $(1)(2) \p \C \p$ Th. \s
\s
\s \s
\s*
6. \s $m,p,q \in \N \p \C \p \frac{p}{q} = \frac{mp}{mq}$. \s (Eucl. VII, 17)
\s
\s \s
\s*
7. \s $p,q \in \N \p m \in \N \p m D p \p m D q : \C \p \frac{p}{q} = \frac{p/m}{q/m}$. \s
\s
\s \s
\s*
8. \s $p,q,p',q' \in \N \p p \pi q \p p' \pi q' \p \frac{p}{q} = \frac{p'}{q'} : \C : p = p' \p q =q'$. \s
\s
\s \s
\s*
9. \s $p,q,p',q' \in \N \p p' \pi q' \p \frac{p}{q} = \frac{p'}{q'} : \C : p' / p = q' / q = M \such D (p,q)$. \s
\s
\s \s
\s*
10. \s $p,q,p',q' \in \N \p \frac{p}{q} = \frac{p'}{q'} \p p \pi q \p q' \smaller q : = \abs$. \s (Eucl. VII, 21)
\s
\s \s
\s*
11. \s $p,q,p',q' \in \N : \C : \frac{p}{q} = \frac{p'}{q'} \p = \p \frac{p}{p'} = \frac{q}{q'} \p = \p \frac{q}{p} = \frac{q'}{p'}$. \s (Eucl. VII, 13)
\s
\s \s
\s*
12. \s $p,q \in \N \p \C :: [m \in] : m \in \N \p m \frac{p}{q} \in \N \ppp \no = \abs$. \s
\s
\s \s
\s*
12'. \s $a\in \R \p \C :: [m \in] : m \in \N \p ma \in \N \ppp \no = \abs$. \s
\s
\s \s
\end{paracol}

\peanoPage{14} % page-number 14

\columnratio{0.1, 0.3, 0.1, 0.1, 0.3, 0.1}
\begin{paracol}{6}
\raggedright
13. \s $p,q,p',q' \in \N \p \C :: [(r,s,l) \in] : r,s,t \in \N \p \frac{p}{q} = \frac{r}{t} \p \frac{p'}{q'} = \frac{s}{t} \ppp \no = \abs$. \s
\s
\s \s
\s*
13'. \s $a,b \in \R \p \C :: [(r,s,t) \in] : r,s,t \in \N \p a = \frac{r}{t} \p b = \frac{s}{t} \ppp \no = \abs$. \s
\s
\s \s
\s*
14. \s $a,b,c \in \R \p \C :: [(m,n,p,q) \in] : m,n,p,q \in \N \p a = \frac{m}{q} \p b = \frac{n}{q} \p c = \frac{p}{q} \ppp \no = \abs$. \s
\s
\s \s
\s*
15. \s $p,q,r \in \N \p a = \frac{p}{r} \p b = \frac{q}{r} : \C : a = b \p = \p p = q$. \s
\s
\s \s
\s*
16. \s $m \in \N \p a,b \in \R \p a = b \p ma \in \N : \C \p mb \in N$. \s
\s
\s \s
\s*
17. \s $a,b,c \in \R \p \C \ppp a = a$. \s
\s
\s \s
\s*
\s $\hspace{1.72cm} \C \ppp a = b \p = \p b = a$. \s
\s
\s \s
\s*
\s $\hspace{1.72cm} \C \ppp a = b \p b = c : \C \p a = c$. \s
\s
\s \s
\s*
18. \s $N \C \R$. \s
\s
\s \s
\end{paracol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Definitions}
\peanoHeadingSmall{Definitiones.}
\s
\peanoHeadingSmall{Definitions.}
\end{translateTwoCol}

\columnratio{0.1, 0.3, 0.1, 0.1, 0.3, 0.1}
\begin{paracol}{6}
\raggedright
19. \s $a,b \in \R \p \C :: a \smaller b \p = \ppp x \in \N \p xa, xb \in \N : \C \p xa < xb$. \s
\s
\s \s
\s*
20. \s $a,b \in \R \p \C : b \larger a \p = \p a \smaller b$. \s
\s
\s \s
\end{paracol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Theorems}
\peanoHeadingSmall{Theoremata.}
\s
\peanoHeadingSmall{Theorems.}
\end{translateTwoCol}

\columnratio{0.1, 0.3, 0.1, 0.1, 0.3, 0.1}
\begin{paracol}{6}
\raggedright
21. \s $p,q,r \in \N \p a = \frac{p}{r} \p b = \frac{q}{r} : \C : a \smaller b \p = \p p \smaller q$. \s
\s
\s \s
\s*
22. \s $p,q,p',q' \in \N \p \C : \frac{p}{q} \smaller \frac{p'}{q'} \p = \p pq' \smaller p'q$. \s
\s
\s \s
\s*
23. \s $p,q,r \in \N \p a = \frac{r}{p} \p b = \frac{r}{q} : \C : a \smaller b \p = \p p > q$. \s
\s
\s \s
\s*
24. \s $p,q,p',q' \in \N \p \frac{p}{q} \smaller \frac{p'}{q'} : \C \p \frac{p}{q} \smaller \frac{p + p'}{q + q'} \smaller \frac{p'}{q'}$. \s
\s
\s \s
\s*
25. \s $a \in \R \p \C \ppp ? \p \such \larger a : \no = \abs$. \s
\s
\s \s
\s*
26. \s $a \in \R \p \C \ppp \R \p \such \smaller a : \no = \abs$. \s
\s
\s \s
\s*
27. \s $a,b \in \R \p a \smaller b : \C \ppp \R \p \such \larger a \p \such \smaller b : \no = \abs$. \s
\s
\s \s
\s*
28. \s $a,b \in \R : \C \ppp a \smaller b \p a = b : = \abs$. \s
\s
\s \s
\s*
\s $\hspace{1.5cm} \C \ppp a \larger b \p a = b : = \abs$. \s
\s
\s \s
\s*
\s $\hspace{1.5cm} \C \ppp a \smaller b \p a \larger b : = \abs$. \s
\s
\s \s
\s*
\s $\hspace{1.5cm} \C \ppp a \no \smaller b \p a \no = b \p a \no \larger b : = \abs$. \s
\s
\s \s
\s*
29. \s $a,b,c \in \R : \C \ppp a \smaller \cup = b \p b \smaller c : \C : a \smaller c$. \s
\s
\s \s
\s*
\s $\hspace{1.7cm} \C \ppp a \smaller b \p b \smaller \cup = c : \C : a \smaller c$. \s
\end{paracol}

\peanoPage{15} % page-number 15

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Definitions}
\peanoHeadingSmall{Definitiones.}
\s
\peanoHeadingSmall{Definitions.}
\end{translateTwoCol}

\columnratio{0.1, 0.3, 0.1, 0.1, 0.3, 0.1}
\begin{paracol}{6}
\raggedright
30. \s $a,b \in \R \p \C \p a + b = [c \in] (c \in  \R \ppp x \in \N \p x a, x b, x c \in \N : \C_x \p xa + xb = xc)$. \s
\s
\s \s
\s*
31. \s $a,b \in \R \p \C :: b - a = \ppp [x \in] (x \in \R \p a + x = b)$. \s
\s
\s \s
\s*
32. \s $a,b \in \R \p \C \p ab = [c \in] (c \in \R \ppp x \in \N \p xa, (xa)b, xc \in \N : \C_x \p (xa) b = xc)$. \s
\s
\s \s
\s*
33. \s $a,b \in \R \p \C \p b / a = [x \in] (x \in \R \p ax = b)$. \s
\s
\s \s
\end{paracol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Theorems}
\peanoHeadingSmall{Theoremata.}
\s
\peanoHeadingSmall{Theorems.}
\end{translateTwoCol}

\columnratio{0.1, 0.3, 0.1, 0.1, 0.3, 0.1}
\begin{paracol}{6}
\raggedright
34. \s $p,q,r \in \N \p \C \frac{p}{r} + \frac{q}{r} = \frac{p + q}{r}$. \s
\s
\s \s
\s*
35. \s $a,b \in \R \p \C \p a + b \in \R$. \s
\s
\s \s
\s*
36. \s $p,q,r \in \N \p p \smaller q : \C \p \frac{q}{r} - \frac{p}{r} = \frac{q-p}{r}$. \s
\s
\s \s
\s*
37. \s $a,b \in \R \p a \smaller b : \C \p b - a \in \R$. \s
\s
\s \s
\s*
38. \s $p,q,p',q' \in \N \p \C \p \frac{p}{q} \frac{p'}{q'} = \frac{pp'}{qq'}$. \s
\s
\s \s
\s*
39. \s $a,b \in \R \p \C \p ab \in \R$. \s
\s
\s \s
\s*
40. \s $p,q,p',q' \in \N \p \C \p \frac{p}{q} / \frac{p'}{q'} = \frac{pq'}{p'q}$. \s
\s
\s \s
\s*
41. \s $a,b \in \R \p \C \p b / a \in \R$. \s
\s
\s \s
\s*
42. \s $p,q \in \N \p \C \p \frac{p}{q} = \frac{p}{q}$. \s
\s
\s \s
\end{paracol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{section}{\S9. The system of rationals. Irrationals}
\peanoHeadingMedium{\S9. Rationalum systemata. Irrationales.}
\s
\peanoHeadingMedium{\S9. The system of rationals. Irrationals.}
\s*
\phantomsection
\addcontentsline{toc}{subsection}{Explanation}
\peanoHeadingSmall{Explicatio.}
\s
\peanoHeadingSmall{Explanation.}
\end{translateTwoCol}

\begin{translateTwoCol}
\raggedright
Si $a \in \K \R$, signum $\T a$ legitur \emph{terminus summus}, vel \emph{limes summus classis} $a$. Supra hoc novum ens relationes ac operationes tantum definimus.
\s
If $a \in \K \R$, the sign $\T$ a is read \emph{upper boundary}, or \emph{upper limit of the class} $a$. We shall define only a few relations and operations on this new entity.
\end{translateTwoCol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Definitions}
\peanoHeadingSmall{Definitiones.}
\s
\peanoHeadingSmall{Definitions.}
\end{translateTwoCol}

\columnratio{0.1, 0.3, 0.1, 0.1, 0.3, 0.1}
\begin{paracol}{6}
\raggedright
1. \s $a \in \K \R \p x \in \R : \C :: x \smaller \T a \p = \ppp a \p \such \larger x : \no = \abs$. \s
\s
\s \s
\s*
2. \s $a \in \K \R \p x \in \R : \C ::: x = \T a \p = : \p : a \p \such \larger x : = \abs :: u \in \R \p u \smaller x : \C_x \ppp a \p \such \larger u : \no = \abs$. \s
\s
\s \s
\s*
3. \s $a \in \K \R \p x \in \R : \C \ppp x \larger \T a \p = : x -\no \smaller \T a \p x \no = \T a$. \s
\s
\s \s
\end{paracol}

\peanoPage{16} % page-number 16

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Theorem}
\peanoHeadingSmall{Theorema.}
\s
\peanoHeadingSmall{Theorem.}
\end{translateTwoCol}

\columnratio{0.1, 0.3, 0.1, 0.1, 0.3, 0.1}
\begin{paracol}{6}
\raggedright
4. \s $x \in \R \p \C :: x = \ppp \T : \R \p \such \smaller x$. \s
\s
\s \s
\end{paracol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Explanation}
\peanoHeadingSmall{Explicatio.}
\s
\peanoHeadingSmall{Explanation.}
\end{translateTwoCol}

\begin{translateTwoCol}
\raggedright
Signum $\Q$ legitur \emph{quantitas}, numerosque indicat reales positivos, rationales aut irrationales, $0$ et $\infty$ exceptis.
\s
The sign $\Q$ is read \emph{quantity}, and indicates the positive real numbers, rational or irrational, with the exception of $0$ and $\infty$.
\end{translateTwoCol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Definitions}
\peanoHeadingSmall{Definitiones.}
\s
\peanoHeadingSmall{Definitions.}
\end{translateTwoCol}

\columnratio{0.1, 0.3, 0.1, 0.1, 0.3, 0.1}
\begin{paracol}{6}
\raggedright
5. \s $\Q = [x \in] (a \in K \R : a \no = \abs : \R \in \larger \T a \p \min = \abs : \T a = x \ppp \no = \abs)$. \s
\s
\s \s
\s*
6. \s $a,b \in \Q \p \C :: a = b \p = \ppp \R \p \such \smaller a : = : \R \p \such \smaller b$. \s
\s
\s \s
\s*
7. \s $a,b \in \Q \p \C :: a \smaller b \p = \ppp \R \p \such \larger a \p \such \smaller b : = \no = \abs$. \s
\s
\s \s
\s*
8. \s $a,b \in \Q \p \C : b \larger a \p = \p a \smaller b$. \s
\s
\s \s
\end{paracol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Theorems}
\peanoHeadingSmall{Theoremata.}
\s
\peanoHeadingSmall{Theorems.}
\end{translateTwoCol}

\columnratio{0.1, 0.3, 0.1, 0.1, 0.3, 0.1}
\begin{paracol}{6}
\raggedright
9. \s $a \in \Q \p \C \ppp \R \p \such \smaller a : \no = \abs$. \s
\s
\s \s
\s*
10. \s $a \in \Q \p \C \ppp \R \p \such \larger a : \no = \abs$. \s
\s
\s \s
\s*
11. \s $\R \C \Q$. \s
\s
\s \s
\s*
\vspace{1em}
\s Subsistunt quoque propositiones quae a P17, 28, 29 in \S8 obtinentur, si loco $\R$ legatur $\Q$. \s
\s
\vspace{1em}
\s The propositions obtained from P17, 28, 29 in \S8 also hold, by reading $\Q$ for $\R$. \s
\end{paracol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Definitions}
\peanoHeadingSmall{Definitiones.}
\s
\peanoHeadingSmall{Definitions.}
\end{translateTwoCol}

\columnratio{0.1, 0.3, 0.1, 0.1, 0.3, 0.1}
\begin{paracol}{6}
\raggedright
12. \s $a,b \in Q \p \C \p a + b = T [z \in] ([(x,y) \in] : x,y \in \R \p x \smaller a \p y \smaller b \p x + y = z \ppp \no = \abs)$. \s
\s
\s \s
\s*
13. \s $a,b \in Q \p \C \p ab = T [z \in] ([x,y) \in] : x,y \in \R \p x \smaller a \p y \smaller b \p xy = z \ppp \no = \abs)$. \s
\s
\s \s
\s*
\vspace{1em}
\s Ut valeant hae definitiones, demonstrandum est subsistere propositiones 12 et 13, si $a,b \in \R$. \s
\s
\vspace{1em}
\s In order for these definitions to have meaning, it must be proved that propositions 12 and 13 hold, if $a,b \in \R$. \s
\s*
\vspace{1em}
\s Substractionem et divisionem ut operationes inversas additiones et multiplicationis definire licet, illarumque proprietas demonstrare. \s
\s
\vspace{1em}
\s Subtraction and division could be defined as the inverse operations to addition and multiplication, and their properties could be proved. \s
\end{paracol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{section}{\S10. System of quantities}
\peanoHeadingMedium{\S10. Quantitatum systemata.}
\s
\peanoHeadingMedium{\S10. System of quantities.}
\s*
\phantomsection
\addcontentsline{toc}{subsection}{Explanations}
\peanoHeadingSmall{Explicationes.}
\s
\peanoHeadingSmall{Explanations.}
\end{translateTwoCol}

\begin{translateTwoCol}
\raggedright
Si $a \in \K \Q$, signa $\I a, \E a, \Lfat a$ leguntur: \emph{interior, exterior, limes classis} $a$.
\s
If $a \in \K \Q$, the signs $\I a, \E a, \Lfat a$ are read: \emph{interior, exterior, limit of the class} $a$.
\end{translateTwoCol}

\peanoPage{17} % page-number 17

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Definitions}
\peanoHeadingSmall{Definitiones.}
\s
\peanoHeadingSmall{Definitions.}
\end{translateTwoCol}

\columnratio{0.1, 0.3, 0.1, 0.1, 0.3, 0.1}
\begin{paracol}{6}
\raggedright
1. \s $a \in \K \Q \p \C \I a = \Q [x \in] ([u,v) \in] :: u,v \in \Q \ppp u \smaller x \smaller v \ppp \such \larger u \p \such \smaller v : \C : a : \p : \no = \abs)$. \s
\s
\s \s
\s*
2. \s $a \in \K \Q \p \C \p \E a = \I (\no a)$. \s
\s
\s \s
\s*
3. \s $a \in \K \Q \p \C \p \Lfat a = (\no \I a) (\no \E a)$. \s
\s
\s \s
\end{paracol}

\begin{translateTwoCol}
\centering
\phantomsection
\addcontentsline{toc}{subsection}{Theorems}
\peanoHeadingSmall{Theoremata.}
\s
\peanoHeadingSmall{Theorems.}
\end{translateTwoCol}

\columnratio{0.1, 0.3, 0.1, 0.1, 0.3, 0.1}
\begin{paracol}{6}
\raggedright
4. \s $a \in \K \Q \p x,u,v \in Q \p u \smaller x \smaller v \p (\such \larger u \p \such \smaller v : \C a) : \C \p x \in \I a$. \s
\s
\s \s
\s*
5. \s $a \in \K \Q \p x \in \I a : \C : [(u,v) \in] (u,v \in \Q \ppp u \smaller x \smaller v \ppp \such \larger u \p \such \smaller v : \C : a) \no = \abs$. \s
\s
\s \s
\s*
\emph{Dem.} \s P1 $=$ (P4)(P5). \s
\s
\s \s
\s*
6. \s $a \in \K \Q \p u,v \in \Q \p (\such \larger u \p \such \smaller v : \C a) \ppp \C \ppp \such \larger u \p \such \smaller v : \C \I a$. \s
\s
\s \s
\s*
\emph{Dem.} \s P6 $=$ P4. \s
\s
\s \s
\s*
7. \s $a \in \K \Q \p \C \p \I a \C a$. \s
\s
\s \s
\s*
8. \s $ a \in \K \Q \p \C \p \II a = \I a$. \s
\s
\s \s
\s*
\emph{Dem.} \s Hp. $(\I a) [a]$ P7 $: \C \p \II a \C \I a$ \s (1)
\s
\s \s
\s*
\s Hp. $x,u,v \in \Q \p u \smaller x \smaller v \p (\such \larger u \p \such \smaller v : \C a) \p$ P6 $: \C : u,v \in \Q \p u \smaller x \smaller v \p (\such \larger u \p \such \smaller v : \C \I a)$ \s (2)
\s
\s \s
\s*
\s Hp. $x \in \I a \p (2) : \C : x \in \II a$ \s (3)
\s
\s \s
\s*
\s Hp. $(3) : \C : \I a \II a$ \s (4)
\s
\s \s
\s*
\s Hp. $(1) \p (4) : \C :$ Ts. \s (Theor.)
\s
\s \s
\s*
9. \s $a,b \in \K \Q \p a \C b : \C \p \I a \C \I b$ \s
\s
\s \s
\s*
\emph{Dem.} \s Hp. $x,u,v \in \Q \p u \smaller x \smaller v \p (\such \larger u \p \such \smaller v : \C a) : \C \ppp \such \larger u \p \such \smaller v : \C b$ \s (1)
\s
\s \s
\s*
\s Hp. $x \in \I a : \C : x \in \I b$ \s (Theor.)
\s
\s \s
\s*
10. \s $a,b \in \K \Q : \C : \I (ab) \C \I a$ \s
\s
\s \s
\s*
\emph{Dem.} \s $(ab, a) [a,b]$ P9 $\p = \p$ P10 \s
\s
\s \s
\s*
11. \s $a,b \in \K \Q \p \C \p \I (ab) \C (\I a) (\I b)$ \s
\s
\s \s
\s*
\emph{Dem.} \s P11 $= :$ P10 $\p \cap \p (b,a) [a,b]$ P10 \s
\s
\s \s
\s*
12. \s $a,b \in \K \Q \p \C \p \I a \C \I (a \cup b)$ \s
\s
\s \s
\s*
13. \s $a,b \in \K \Q \p \C \p \I a \cup \I b \C \I (a \cup b)$ \s
\s
\s \s
\s*
14. \s $a,b \in \K \Q \p \C \p \I (ab) = (\I a)(\I b)$ \s
\s
\s \s
\s*
\emph{Dem.} \s Hp. P11 $: \C \p \I (ab) \C (\I a)(\I b)$ \s (1)
\s
\s \s
\end{paracol}

\peanoPage{18} % page-number 18

\columnratio{0.1, 0.3, 0.1, 0.1, 0.3, 0.1}
\begin{paracol}{6}
\raggedright
\s Hp. $x \in \Q \p u,v \in \Q \p u \smaller x \smaller v \p (\such \larger u \p \such \smaller v : \C a) \p u',v' \in \Q \p u' \smaller x \smaller v' \p (\such \larger u' \p \such \smaller v' : \C b) \p u'' = M (u \cup u') \p v'' = \mini (v,v') : \C : u'', v'' \in Q \p u '' \smaller x \smaller v'' \p (\such \larger u'' \p \such \larger v'' : \C : ab)$ \s (2)
\s
\s \s
\s*
\s Hp. $x \in \I a \p x \in \I b \p (2) : \C \p x \in \I (ab)$ \s (3)
\s
\s \s
\s*
\s Hp. $(3) : \C : (\I a)(\I b) \C \I (ab)$ \s (4)
\s
\s \s
\s*
\s Hp. $(1)\p (4) : \C \p$ Ts. \s
\s
\s \s
\s*
15. \s $a \in \K \Q \p \C \p \E a \C - a$ \s
\s
\s \s
\s*
\emph{Dem.} \s P15 $= (\no a) [a]$ P7 \s
\s
\s \s
\s*
16. \s $a \in \K \Q \p \C \ppp \I a \p \E a : = \abs$ \s
\s
\s \s
\s*
\emph{Dem.} \s Hp. P7 $.$ P15 $: \C \ppp \I a \p \E a : \C : a \no a : = \abs$ \s
\s
\s \s
\s*
17. \s $a \in \K \Q \p \C \p \I \E a = \E a$ \s
\s
\s \s
\s*
\emph{Dem.} \s P17 $= (\no a) [a]$ P8 \s
\s
\s \s
\s*
18. \s $a,b \in \K \Q \p b \C a : \C \p \E a \C \E b$ \s
\s
\s \s
\s*
\emph{Dem.} \s P18 $= (\no a, \no b) [a,b]$ P9 \s
\s
\s \s
\s*
19. \s $a,b \in \K \Q \p \C : \E a \cup \E b \p \C \E (ab)$ \s
\s
\s \s
\s*
20. \s $a,b \in \K \Q \p \C \p \E (a \cup b) = (\E a) (\E b)$ \s
\s
\s \s
\s*
\emph{Dem.} \s P20 $= (\no a, \no b) [a,b]$ P14 \s
\s
\s \s
\s*
21. \s $a \in \K \Q \p \C \p \Lfat (\no a) = \Lfat a$ \s
\s
\s \s
\s*
22. \s $a \in \K \Q \p \C \ppp \I a \p \Lfat a : = \abs$ \s
\s
\s \s
\s*
\s $\hspace{1.68cm} \C \ppp \E a \p \Lfat a : = \abs$ \s
\s
\s \s
\s* 
\s $\hspace{1.68cm} \C \ppp \no \I a \p \no \E a \p \no \Lfat a : = \abs$ \s
\s
\s \s
\s*
\emph{Dem.} \s P22 $=$ P3 \s
\s
\s \s
\s*
23. \s $a \in \K \Q \p \C : a \C \p \I a \cup \Lfat a$ \s
\s
\s \s
\s*
24. \s $a \in \K \Q \p \C \p \I (a \Lfat a) = \abs$ \s
\s
\s \s
\s*
\emph{Dem.} \s Hp. P14 $\p$ P7 $\p$ P22 $: \C : \I (a \Lfat a) \p = \p \I a \I \Lfat a \p \C \p \I a \Lfat a \p = \p \abs$ \s
\s
\s \s
\s*
25. \s $a,b \in \K \Q \p a \C b : \C : \Lfat a \p \C \p \I b \cup \Lfat b$ \s
\s
\s \s
\s*
\emph{Dem.} \s Hp. P18 $: \C : \E b \C \E a : \C : \I a \cup \Lfat a \p \C \p \I b \cup \Lfat b : \C \p$ Ts. \s
\s
\s \s
\s*
26. \s $a,b \in \K \Q \p \C : \Lfat (ab) \C \p \I a \Lfat b \cup \I b \Lfat a \cup \Lfat a \Lfat b$ \s
\s
\s \s
\s*
\emph{Dem.} \s Hp. $\C : ab \C a \p ab \C b \p$ P25 $: \C : \Lfat (ab) \C \I a \cup \Lfat a \p \Lfat (ab) \C \I b \cup \Lfat b : \C : \Lfat (ab) \C (\I a \cup \Lfat a) (\I b \cup \Lfat b) \p \Lfat (ab) (\I a) (\I b) = \Lfat (ab) \I (ab) = \abs : \C :$ Ts. \s
\s
\s \s
\s*
26' \s $a,b \in \K \Q \p \C \p \Lfat (ab) \C \Lfat a \cup \Lfat b$ \s
\s
\s \s
\s*
27. \s $a,b \in \K \Q \p \C : \Lfat (a \cup b) = \Lfat a \E b \cup \Lfat b \E a \cup \Lfat a \Lfat b $ \s
\s
\s \s
\s*
\emph{Dem.} \s P27 $= (\no a, \no b) [a,b]$ P26 \s
\s
\s \s
\end{paracol}

\peanoPage{19} % page-number 19

\columnratio{0.1, 0.3, 0.1, 0.1, 0.3, 0.1}
\begin{paracol}{6}
\raggedright
27'. \s $a,b \in \K \Q \p \C : \Lfat (a \cup b) \C \Lfat a \cup \Lfat b$ \s
\s
\s \s
\s*
28. \s $a \in \K \Q \p \C \p \Lfat \I a \C \Lfat a$ \s
\s
\s \s
\s*
\emph{Dem.} \s Hp. P7 $: \C : \I a \C a \p$ P25 $: \C : \Lfat \I a \C \I a \cup \Lfat a$ \s (1)
\s
\s \s
\s*
\s Hp. P8 $\p$ P22 $: \C \p \Lfat \I a \I a = \Lfat \I a \II a = \abs$ \s (2)
\s
\s \s
\s*
\s $(1)(2) \p \C \p$ Theor. \s
\s
\s \s
\s*
28'. \s $a \in \K \Q \p \C \Lfat \E a \C \Lfat a$ \s
\s
\s \s
\s*
29. \s $a \in \K \Q \p \C \p \LfatLfat a \C \Lfat \I a \cup \Lfat \E a$ \s
\s
\s \s
\s*
\emph{Dem.} \s Hp. $C : \LfatLfat a = \Lfat (\I a \cup \E a) \p$ P27' $: \C \p$ Ts. \s
\s
\s \s
\s*
29'. \s $a \in \K \Q \p \C \p \LfatLfat a \C \Lfat a$ \s
\s
\s \s
\s*
\emph{Dem.} \s P29 $\p$ P28 $\p$ P28' $: \C \p$ Theor. \s
\s
\s \s
\s*
30. \s $a \in \K \Q \p \C \p \Lfat a = \I\Lfat a \cup \LfatLfat a$ \s
\s
\s \s
\s*
\emph{Dem.} \s Hp. P23 $: \C \p \Lfat a \C \I\Lfat a \cup \LfatLfat a$ \s (1)
\s
\s \s
\s*
\s Hp. P7 $: \C \p \I \Lfat a \C \Lfat a$ \s (2)
\s
\s \s
\s*
\s Hp. P29' $: \C \p \LfatLfat a \C \Lfat a$ \s (3)
\s
\s \s
\s*
\s $(1)(2)(3) \p \C \p$ Theor. \s
\s
\s \s
\s*
31. \s $a \in \K \Q \p \C \p \Lfat \I \Lfat a \C \LfatLfat a$\s
\s
\s \s
\s*
\emph{Dem.} \s P31 $= (\Lfat a) [a]$ P28 \s
\s
\s \s
\s*
32. \s $a \in \K \Q \p \C \p \I \LfatLfat a = \abs$ \s
\s
\s \s
\s*
\emph{Dem.} \s Hp. P29' $: \C : \LfatLfat a = \Lfat a \LfatLfat a \p (\Lfat a) [a]$ P24 $: \C$ Ts. \s
\s
\s \s
\s*
33. \s $a \in \K \Q \p \C : \I\Lfat\I\Lfat a = \abs$ \s
\s
\s \s
\s*
\emph{Dem.} \s P31 $\p$ P32 $: \C \p P33$ \s
\s
\s \s
\s*
34. \s $a \in \K \Q \p \C \p \LfatLfat\Lfat a = \LfatLfat a$ \s
\s
\s \s
\s*
\emph{Dem.} \s $(\Lfat a) [a]$ P30 $/p$ P32 $: \C \p$ Theor. \s
\s
\s \s
\s*
35. \s $a,b \in \K \Q \p \C \p \I a \Lfat b \C \Lfat (ab)$ \s
\s
\s \s
\s*
\emph{Dem.} \s Hp. P14 $: \C \p \I a \Lfat b \I (ab) = \I a \I b \Lfat b = \abs$ \s (1)
\s
\s \s (1)
\s*
\s Hp. P2 $\p$ P14 $: \C \p \I a \Lfat b \E (ab) = \I a \Lfat b \I (\no a \cup \no b) = \I (a \no b) \Lfat b = \I a \E b \Lfat b = \abs$ \s (2)
\s
\s \s (2)
\s*
\s $(1)(2) \C$ Theor. \s
\s
\s \s
\s*
36. \s $a,b \in \K \Q \p \C \p \I a \Lfat b \cup \I b \Lfat a \C \Lfat ab$. \s (Vide P26)
\s
\s \s (Cf. p. 26)
\s*
\emph{Dem.} \s P36 $= :$ P35 $\p (b,a) [a,b]$ P35 \s
\s
\s \s
\s*
37. \s $a,b \in \K \Q \p \C \p \E a \Lfat b \cup \E b \Lfat a \cup \Lfat (a \cup b)$. \s (Vide P27)
\s
\s \s (Cf. p. 27)
\s*
\emph{Dem.} \s P37 $= (\no a, \no b)[a,b]$ P36 \s
\s
\s \s
\s*
38. \s $a,b \in \K \Q \p \C \p \I (a \cup b) \C \I a \cup \I b \cup \Lfat a \Lfat b$ \s (Vide P13)
\s
\s \s (Cf. p. 13)
\s*
\emph{Dem.} \s Hp. $\C \p \I (a \cup b) \C (\I a \cup \Lfat a \cup \E a) (\I b \cup \Lfat b \cup \E b)$ \s (1)
\s
\s \s  (1)
\s*
\s Hp. P20 $\p$ P16 $: \C \p \I (a \cup b) \E a \E b = \I (a \cup b) \E (a \cup b) = \abs$ \s (2)
\s
\s \s (2)
\end{paracol}

\peanoPage{20} % page-number 20

\columnratio{0.1, 0.3, 0.1, 0.1, 0.3, 0.1}
\begin{paracol}{6}
\raggedright
\s Hp. P37$: \C : \I (a \cup b)(\E a \Lfat b \cup \E b \Lfat a) \p \C \p \I (a \cup b) \Lfat (a \cup b) \p = \abs$ \s (3)
\s
\s \s (3)
\s*
\s $(1)(2)(3) \p \C \p$ Theor. \s
\s
\s \s
\s*
38'. \s $a,b \in \K \Q \p \C \p \E (ab) \C \E a \cup \E b \cup \Lfat a \Lfat b$ \s (Vide P19)
\s
\s \s (Cf. p. 19)
\s*
39. \s $a \in \K \Q \p \C \p \I\Lfat a \Lfat\I a = \abs$ \s
\s
\s \s
\s*
\emph{Dem.} \s Hp. P36 $: \C : \I\Lfat a \Lfat\I a \C \Lfat (\Lfat a \I a) = \abs$ \s
\s
\s \s
\s*
40. \s $a \in \K \Q \p \C \p \Lfat \I a \C \LfatLfat a$ \s
\s
\s \s
\s*
\emph{Dem.} \s Hp. P28 $\p$ P30 $\p$ P39 $: \C$ Theor. \s
\s
\s \s
\s*
40'. \s $a \in \K \Q \p \C \p \Lfat \E a \C \LfatLfat a$ \s
\s
\s \s
\s*
41. \s $a \in \K \Q \p \C \LfatLfat a = \Lfat\I a \cup \Lfat\E a$ \s
\s
\s \s
\s*
\emph{Dem.} \s P29 $\p$ P40 $\p$ P40' $: \C \p$ Theor. \s
\s
\s \s
\s*
42. \s $a \in \K \Q \p \C \p \I\Lfat\I a = \abs$ \s
\s
\s \s
\s*
\s $\hspace{1.68cm} \C \p \I\Lfat\E a = \abs$ \s
\s
\s \s
\s*
\s $\hspace{1.68cm} \C \p \LfatLfat\I a = \Lfat\I a$ \s
\s
\s \s
\s*
\s $\hspace{1.68cm} \C \p \LfatLfat\E a = \Lfat\E a$ \s
\s
\s \s
\s*
43. \s $a,b \in \K \Q \p \C \p \I (\I a \cup \I b) = \I a \cup \I b$ \s
\s
\s \s
\s*
\emph{Dem.} \s Hp. P7 $: \C \p \I (\I a \cup \I b) \C \I a \cup \I b$ \s (1)
\s
\s \s (1)
\s*
\s Hp. P8 $\p$ P13 $: \C : \I a \cup \I b \p = \p \II a \cup \II b \p \C \p \I (\I a \cup \I b)$ \s (2)
\s
\s \s (2)
\s*
\s $(1)(2) \C$ Theor. \s
\s
\s \s
\s*
44. \s $a,b \in \K \Q \p \C \p \I (\LfatLfat a \cup \LfatLfat b) = \abs$ \s
\s
\s \s
\s*
\emph{Dem.} \s Hp. P38 $\p$ P32 $\p$ P34 $: \C \p \I (\LfatLfat a \cup \LfatLfat b) \C \LfatLfat a \LfatLfat b \C \LfatLfat a$ \s (1)
\s
\s \s (1)
\s*
\s Hp. $(1) \p$ P8 $: \C \p \I (\LfatLfat a \cup \LfatLfat b) \cup \I \LfatLfat a = \abs$ \s
\s
\s \s
\s*
45. \s $a \in \K \Q \p \C \p \I (\I a \cup \E a) = \I a \cup \E a$ \s
\s
\s \s
\s*
\emph{Dem.} \s P8 $\p$ P17 $\p (\no a) [b]$ P43 $: \C \p$ Theor. \s
\s
\s \s
\s*
45'. \s $a \in \K \Q \p \C \p \E\Lfat a = \I a \cup \E a$ \s
\s
\s \s
\s*
46. \s $a \in \K \Q \p \C \p \E\I a = \no (\I a \cup \Lfat\I a)$ \s
\s
\s \s
\s*
46'. \s $a \in \K \Q \p \C \p \EE a = \no (\E a \cup \Lfat\E a)$ \s
\s
46'. \s \s
\end{paracol}

\peanoPage{END} 


%\vspace{1em}
%\columnratio{1} \begin{paracol}{1} \centering \hdashrule{\columnwidth}{0.1mm}{0.1mm 1mm} \end{paracol} % end of page-numbering

\phantomsection
\addcontentsline{toc}{chapter}{Endnotes}
\vspace{1em}
\eendnoteHeading{\textbf{Endnotes}}
\theendnotes
\end{document}
